### 2016-03-29 RMS/corrections.tex
modified version wrt Zhou Min's comments

\documentclass[10pt,journal]{IEEEtran}
%
% If IEEEtran.cls has not been installed into the LaTeX system files,
% manually specify the path to it like:
% \documentclass[10pt,journal,compsoc]{../sty/IEEEtran}

% Some very useful LaTeX packages include:
% (uncomment the ones you want to load)

% *** MISC UTILITY PACKAGES ***
%
%\usepackage{ifpdf}
% Heiko Oberdiek's ifpdf.sty is very useful if you need conditional
% compilation based on whether the output is pdf or dvi.
% usage:
% \ifpdf
%   % pdf code
% \else
%   % dvi code
% \fi
% The latest version of ifpdf.sty can be obtained from:
% http://www.ctan.org/tex-archive/macros/latex/contrib/oberdiek/
% Also, note that IEEEtran.cls V1.7 and later provides a builtin
% \ifCLASSINFOpdf conditional that works the same way.
% When switching from latex to pdflatex and vice-versa, the compiler may
% have to be run twice to clear warning/error messages.

% *** CITATION PACKAGES ***
%
\ifCLASSOPTIONcompsoc
  % IEEE Computer Society needs nocompress option
  % requires cite.sty v4.0 or later (November 2003)
  \usepackage[nocompress,noadjust]{cite}
\else
  % normal IEEE
  \usepackage[noadjust]{cite}
\fi
% cite.sty was written by Donald Arseneau
% V1.6 and later of IEEEtran pre-defines the format of the cite.sty package
% \cite{} output to follow that of IEEE. Loading the cite package will
% result in citation numbers being automatically sorted and properly
% "compressed/ranged". e.g., [1], [9], [2], [7], [5], [6] without using
% cite.sty will become [1], [2], [5]--[7], [9] using cite.sty. cite.sty's
% \cite will automatically add leading space, if needed. Use cite.sty's
% noadjust option (cite.sty V3.8 and later) if you want to turn this off
% such as if a citation ever needs to be enclosed in parenthesis.
% cite.sty is already installed on most LaTeX systems. Be sure and use
% version 5.0 (2009-03-20) and later if using hyperref.sty.
% The latest version can be obtained at:
% http://www.ctan.org/tex-archive/macros/latex/contrib/cite/
% The documentation is contained in the cite.sty file itself.
%
% Note that some packages require special options to format as the Computer
% Society requires. In particular, Computer Society  papers do not use
% compressed citation ranges as is done in typical IEEE papers
% (e.g., [1]-[4]). Instead, they list every citation separately in order
% (e.g., [1], [2], [3], [4]). To get the latter we need to load the cite
% package with the nocompress option which is supported by cite.sty v4.0
% and later. Note also the use of a CLASSOPTION conditional provided by
% IEEEtran.cls V1.7 and later.

% *** GRAPHICS RELATED PACKAGES ***
%
\ifCLASSINFOpdf
  % \usepackage[pdftex]{graphicx}
  % declare the path(s) where your graphic files are
  % \graphicspath{{../pdf/}{../jpeg/}}
  % and their extensions so you won't have to specify these with
  % every instance of \includegraphics
  % \DeclareGraphicsExtensions{.pdf,.jpeg,.png}
\else
  % or other class option (dvipsone, dvipdf, if not using dvips). graphicx
  % will default to the driver specified in the system graphics.cfg if no
  % driver is specified.
  % \usepackage[dvips]{graphicx}
  % declare the path(s) where your graphic files are
  % \graphicspath{{../eps/}}
  % and their extensions so you won't have to specify these with
  % every instance of \includegraphics
  % \DeclareGraphicsExtensions{.eps}
\fi
% graphicx was written by David Carlisle and Sebastian Rahtz. It is
% required if you want graphics, photos, etc. graphicx.sty is already
% installed on most LaTeX systems. The latest version and documentation
% can be obtained at: 
% http://www.ctan.org/tex-archive/macros/latex/required/graphics/
% Another good source of documentation is "Using Imported Graphics in
% LaTeX2e" by Keith Reckdahl which can be found at:
% http://www.ctan.org/tex-archive/info/epslatex/
%
% latex, and pdflatex in dvi mode, support graphics in encapsulated
% postscript (.eps) format. pdflatex in pdf mode supports graphics
% in .pdf, .jpeg, .png and .mps (metapost) formats. Users should ensure
% that all non-photo figures use a vector format (.eps, .pdf, .mps) and
% not a bitmapped formats (.jpeg, .png). IEEE frowns on bitmapped formats
% which can result in "jaggedy"/blurry rendering of lines and letters as
% well as large increases in file sizes.
%
% You can find documentation about the pdfTeX application at:
% http://www.tug.org/applications/pdftex

% *** MATH PACKAGES ***
%
\usepackage[cmex10]{amsmath}
% A popular package from the American Mathematical Society that provides
% many useful and powerful commands for dealing with mathematics. If using
% it, be sure to load this package with the cmex10 option to ensure that
% only type 1 fonts will utilized at all point sizes. Without this option,
% it is possible that some math symbols, particularly those within
% footnotes, will be rendered in bitmap form which will result in a
% document that can not be IEEE Xplore compliant!
%
% Also, note that the amsmath package sets \interdisplaylinepenalty to 10000
% thus preventing page breaks from occurring within multiline equations. Use:
%\interdisplaylinepenalty=2500
% after loading amsmath to restore such page breaks as IEEEtran.cls normally
% does. amsmath.sty is already installed on most LaTeX systems. The latest
% version and documentation can be obtained at:
% http://www.ctan.org/tex-archive/macros/latex/required/amslatex/math/

% *** SPECIALIZED LIST PACKAGES ***
%
%\usepackage{algorithmic}
% algorithmic.sty was written by Peter Williams and Rogerio Brito.
% This package provides an algorithmic environment fo describing algorithms.
% You can use the algorithmic environment in-text or within a figure
% environment to provide for a floating algorithm. Do NOT use the algorithm
% floating environment provided by algorithm.sty (by the same authors) or
% algorithm2e.sty (by Christophe Fiorio) as IEEE does not use dedicated
% algorithm float types and packages that provide these will not provide
% correct IEEE style captions. The latest version and documentation of
% algorithmic.sty can be obtained at:
% http://www.ctan.org/tex-archive/macros/latex/contrib/algorithms/
% There is also a support site at:
% http://algorithms.berlios.de/index.html
% Also of interest may be the (relatively newer and more customizable)
% algorithmicx.sty package by Szasz Janos:
% http://www.ctan.org/tex-archive/macros/latex/contrib/algorithmicx/

\usepackage{algpseudocode}

% *** ALIGNMENT PACKAGES ***
%
%\usepackage{array}
% Frank Mittelbach's and David Carlisle's array.sty patches and improves
% the standard LaTeX2e array and tabular environments to provide better
% appearance and additional user controls. As the default LaTeX2e table
% generation code is lacking to the point of almost being broken with
% respect to the quality of the end results, all users are strongly
% advised to use an enhanced (at the very least that provided by array.sty)
% set of table tools. array.sty is already installed on most systems. The
% latest version and documentation can be obtained at:
% http://www.ctan.org/tex-archive/macros/latex/required/tools/

% IEEEtran contains the IEEEeqnarray family of commands that can be used to
% generate multiline equations as well as matrices, tables, etc., of high
% quality.

% *** SUBFIGURE PACKAGES ***
\ifCLASSOPTIONcompsoc
  \usepackage[caption=false,font=footnotesize,labelfont=sf,textfont=sf]{subfig}
\else
  \usepackage[caption=false,font=footnotesize]{subfig}
\fi
% subfig.sty, written by Steven Douglas Cochran, is the modern replacement
% for subfigure.sty, the latter of which is no longer maintained and is
% incompatible with some LaTeX packages including fixltx2e. However,
% subfig.sty requires and automatically loads Axel Sommerfeldt's caption.sty
% which will override IEEEtran.cls' handling of captions and this will result
% in non-IEEE style figure/table captions. To prevent this problem, be sure
% and invoke subfig.sty's "caption=false" package option (available since
% subfig.sty version 1.3, 2005/06/28) as this is will preserve IEEEtran.cls
% handling of captions.
% Note that the Computer Society format requires a sans serif font rather
% than the serif font used in traditional IEEE formatting and thus the need
% to invoke different subfig.sty package options depending on whether
% compsoc mode has been enabled.
%
% The latest version and documentation of subfig.sty can be obtained at:
% http://www.ctan.org/tex-archive/macros/latex/contrib/subfig/

% *** FLOAT PACKAGES ***
%
%\usepackage{fixltx2e}
% fixltx2e, the successor to the earlier fix2col.sty, was written by
% Frank Mittelbach and David Carlisle. This package corrects a few problems
% in the LaTeX2e kernel, the most notable of which is that in current
% LaTeX2e releases, the ordering of single and double column floats is not
% guaranteed to be preserved. Thus, an unpatched LaTeX2e can allow a
% single column figure to be placed prior to an earlier double column
% figure. The latest version and documentation can be found at:
% http://www.ctan.org/tex-archive/macros/latex/base/

%\usepackage{stfloats}
% stfloats.sty was written by Sigitas Tolusis. This package gives LaTeX2e
% the ability to do double column floats at the bottom of the page as well
% as the top. (e.g., "\begin{figure*}[!b]" is not normally possible in
% LaTeX2e). It also provides a command:
%\fnbelowfloat
% to enable the placement of footnotes below bottom floats (the standard
% LaTeX2e kernel puts them above bottom floats). This is an invasive package
% which rewrites many portions of the LaTeX2e float routines. It may not work
% with other packages that modify the LaTeX2e float routines. The latest
% version and documentation can be obtained at:
% http://www.ctan.org/tex-archive/macros/latex/contrib/sttools/
% Do not use the stfloats baselinefloat ability as IEEE does not allow
% \baselineskip to stretch. Authors submitting work to the IEEE should note
% that IEEE rarely uses double column equations and that authors should try
% to avoid such use. Do not be tempted to use the cuted.sty or midfloat.sty
% packages (also by Sigitas Tolusis) as IEEE does not format its papers in
% such ways.
% Do not attempt to use stfloats with fixltx2e as they are incompatible.
% Instead, use Morten Hogholm'a dblfloatfix which combines the features
% of both fixltx2e and stfloats:
%
\usepackage{dblfloatfix}
% The latest version can be found at:
% http://www.ctan.org/tex-archive/macros/latex/contrib/dblfloatfix/

%\ifCLASSOPTIONcaptionsoff
%  \usepackage[nomarkers]{endfloat}
% \let\MYoriglatexcaption\caption
% \renewcommand{\caption}[2][\relax]{\MYoriglatexcaption[#2]{#2}}
%\fi
% endfloat.sty was written by James Darrell McCauley, Jeff Goldberg and 
% Axel Sommerfeldt. This package may be useful when used in conjunction with 
% IEEEtran.cls'  captionsoff option. Some IEEE journals/societies require that
% submissions have lists of figures/tables at the end of the paper and that
% figures/tables without any captions are placed on a page by themselves at
% the end of the document. If needed, the draftcls IEEEtran class option or
% \CLASSINPUTbaselinestretch interface can be used to increase the line
% spacing as well. Be sure and use the nomarkers option of endfloat to
% prevent endfloat from "marking" where the figures would have been placed
% in the text. The two hack lines of code above are a slight modification of
% that suggested by in the endfloat docs (section 8.4.1) to ensure that
% the full captions always appear in the list of figures/tables - even if
% the user used the short optional argument of \caption[]{}.
% IEEE papers do not typically make use of \caption[]'s optional argument,
% so this should not be an issue. A similar trick can be used to disable
% captions of packages such as subfig.sty that lack options to turn off
% the subcaptions:
% For subfig.sty:
% \let\MYorigsubfloat\subfloat
% \renewcommand{\subfloat}[2][\relax]{\MYorigsubfloat[]{#2}}
% However, the above trick will not work if both optional arguments of
% the \subfloat command are used. Furthermore, there needs to be a
% description of each subfigure *somewhere* and endfloat does not add
% subfigure captions to its list of figures. Thus, the best approach is to
% avoid the use of subfigure captions (many IEEE journals avoid them anyway)
% and instead reference/explain all the subfigures within the main caption.
% The latest version of endfloat.sty and its documentation can obtained at:
% http://www.ctan.org/tex-archive/macros/latex/contrib/endfloat/
%
% The IEEEtran \ifCLASSOPTIONcaptionsoff conditional can also be used
% later in the document, say, to conditionally put the References on a 
% page by themselves.

% *** PDF, URL AND HYPERLINK PACKAGES ***
%
%\usepackage{url}
% url.sty was written by Donald Arseneau. It provides better support for
% handling and breaking URLs. url.sty is already installed on most LaTeX
% systems. The latest version and documentation can be obtained at:
% http://www.ctan.org/tex-archive/macros/latex/contrib/url/
% Basically, \url{my_url_here}.

% *** Do not adjust lengths that control margins, column widths, etc. ***
% *** Do not use packages that alter fonts (such as pslatex).         ***
% There should be no need to do such things with IEEEtran.cls V1.6 and later.
% (Unless specifically asked to do so by the journal or conference you plan
% to submit to, of course. )

\usepackage{alltt}
\usepackage{tikz}
\usetikzlibrary{arrows.meta}
\usepackage{multicol}
\newcommand{\hide}[1]{\ignorespaces}
\newcommand{\jx}[1]{{\bf Jiaxiang: }#1{ \bf End}}
\newtheorem{theorem}{Theorem}
\newtheorem{lemma}{Lemma}
\newcommand{\ANSWER}{{\bf RESPONSE: }}
\newcommand{\COMMENT}{{\bf COMMENT: }}

% Define the fontsize in environment {verbatim}
\makeatletter
\def\verbatim{\small\@verbatim \frenchspacing\@vobeyspaces \@xverbatim}
%\def\verbatim@font{\small\ttfamily}
\makeatother

% correct bad hyphenation here
\hyphenation{op-tical net-works semi-conduc-tor}

\begin{document}
%
% paper title
% Titles are generally capitalized except for words such as a, an, and, as,
% at, but, by, for, in, nor, of, on, or, the, to and up, which are usually
% not capitalized unless they are the first or last word of the title.
% Linebreaks \\ can be used within to get better formatting as desired.
% Do not put math or special symbols in the title.
\title{Corrections to Manuscript No. 15-TIE-3480}
%
%
% author names and IEEE memberships
% note positions of commas and nonbreaking spaces ( ~ ) LaTeX will not break
% a structure at a ~ so this keeps an author's name from being broken across
% two lines.
% use \thanks{} to gain access to the first footnote area
% a separate \thanks must be used for each paragraph as LaTeX2e's \thanks
% was not built to handle multiple paragraphs
%
%
%\IEEEcompsocitemizethanks is a special \thanks that produces the bulleted
% lists the Computer Society journals use for "first footnote" author
% affiliations. Use \IEEEcompsocthanksitem which works much like \item
% for each affiliation group. When not in compsoc mode,
% \IEEEcompsocitemizethanks becomes like \thanks and
% \IEEEcompsocthanksitem becomes a line break with idention. This
% facilitates dual compilation, although admittedly the differences in the
% desired content of \author between the different types of papers makes a
% one-size-fits-all approach a daunting prospect. For instance, compsoc 
% journal papers have the author affiliations above the "Manuscript
% received ..."  text while in non-compsoc journals this is reversed. Sigh.

\hide{
\author{Jiaxiang~Liu,
        Min~Zhou,
        Xiaoyu~Song,
        Ming~Gu,
        and~Jiaguang~Sun% <-this % stops a space
\IEEEcompsocitemizethanks{%
\IEEEcompsocthanksitem J. Liu is with the School of Software, Tsinghua
University, Beijing 100084, China, and also with LIX, \'Ecole
Polytechnique, Palaiseau 91120, France.\protect\\ E-mail:
jiaxiang.liu@hotmail.com
\IEEEcompsocthanksitem M. Zhou, M. Gu and J. Sun are with the School
of Software, Tsinghua University, Beijing 100084, China.
\IEEEcompsocthanksitem X. Song is with the Department of Electrical \&
Computer Engineering, Portland State University, Portland, OR
97207-0751, USA.}% <-this % stops an unwanted space
}
%\thanks{Manuscript received April 19, 2005; revised September 17, 2014.}}

% note the % following the last \IEEEmembership and also \thanks - 
% these prevent an unwanted space from occurring between the last author name
% and the end of the author line. i.e., if you had this:
% 
% \author{....lastname \thanks{...} \thanks{...} }
%                     ^------------^------------^----Do not want these spaces!
%
% a space would be appended to the last name and could cause every name on that
% line to be shifted left slightly. This is one of those "LaTeX things". For
% instance, "\textbf{A} \textbf{B}" will typeset as "A B" not "AB". To get
% "AB" then you have to do: "\textbf{A}\textbf{B}"
% \thanks is no different in this regard, so shield the last } of each \thanks
% that ends a line with a % and do not let a space in before the next \thanks.
% Spaces after \IEEEmembership other than the last one are OK (and needed) as
% you are supposed to have spaces between the names. For what it is worth,
% this is a minor point as most people would not even notice if the said evil
% space somehow managed to creep in.
}

% The paper headers
% The only time the second header will appear is for the odd numbered pages
% after the title page when using the twoside option.
% 
% *** Note that you probably will NOT want to include the author's ***
% *** name in the headers of peer review papers.                   ***
% You can use \ifCLASSOPTIONpeerreview for conditional compilation here if
% you desire.

% The publisher's ID mark at the bottom of the page is less important with
% Computer Society journal papers as those publications place the marks
% outside of the main text columns and, therefore, unlike regular IEEE
% journals, the available text space is not reduced by their presence.
% If you want to put a publisher's ID mark on the page you can do it like
% this:
%\IEEEpubid{0000--0000/00\$00.00~\copyright~2014 IEEE}
% or like this to get the Computer Society new two part style.
%\IEEEpubid{\makebox[\columnwidth]{\hfill 0000--0000/00/\$00.00~\copyright~2014 IEEE}%
%\hspace{\columnsep}\makebox[\columnwidth]{Published by the IEEE Computer Society\hfill}}
% Remember, if you use this you must call \IEEEpubidadjcol in the second
% column for its text to clear the IEEEpubid mark (Computer Society jorunal
% papers don't need this extra clearance.)

% use for special paper notices
%\IEEEspecialpapernotice{(Invited Paper)}

% for Computer Society papers, we must declare the abstract and index terms
% PRIOR to the title within the \IEEEtitleabstractindextext IEEEtran
% command as these need to go into the title area created by \maketitle.
% As a general rule, do not put math, special symbols or citations
% in the abstract or keywords.

% make the title area
%\maketitle

% To allow for easy dual compilation without having to reenter the
% abstract/keywords data, the \IEEEtitleabstractindextext text will
% not be used in maketitle, but will appear (i.e., to be "transported")
% here as \IEEEdisplaynontitleabstractindextext when the compsoc 
% or transmag modes are not selected <OR> if conference mode is selected 
% - because all conference papers position the abstract like regular
% papers do.
%\IEEEdisplaynontitleabstractindextext
% \IEEEdisplaynontitleabstractindextext has no effect when using
% compsoc or transmag under a non-conference mode.

% For peer review papers, you can put extra information on the cover
% page as needed:
% \ifCLASSOPTIONpeerreview
% \begin{center} \bfseries EDICS Category: 3-BBND \end{center}
% \fi
%
% For peerreview papers, this IEEEtran command inserts a page break and
% creates the second title. It will be ignored for other modes.
%\IEEEpeerreviewmaketitle

\twocolumn[
  \begin{@twocolumnfalse}
\maketitle

\begin{center}
  \begin{tabular}{r l}
    \hline\hline
    & \\
    Manuscript Number: & 15-TIE-3480\\
    Manuscript Title: & Formal Modeling and Verification of a Rate-Monotonic Scheduling Implementation\\
    & with Real-Time Maude \\
    Submitted to: & Transactions on Industrial Electronics \\
    Manuscript Type: & Regular paper \\
    & \\
    \hline\hline
  \end{tabular}
  \bigskip
\end{center}

  \end{@twocolumnfalse}
]

%\section{Answers to the Reviewers}
\section{Reponses to Reviewer 1}
\subsection{General Comments}
\begin{itemize}

\item
\hide{
The paper shows a realistic implementation of a Rate-Monotonic
Scheduling algorithm using Real-Time Maude (a modeling language based
on rewriting logic) by taking into account the overhead of scheduling
and some other details of the hardware platform. The correctness of
the implementation with respect to the algorithm and the completeness
of the model are verified by model checking, and validated within
different realistic scenarios.}
\COMMENT On the positive side, this paper presents a solid piece of
work. Periodic task scheduling is an important (and difficult)
problem, and the topic of the paper appears to be well suited within
the field of the industrial real-time systems. Overall the paper is
well structured, technically accurate, presents the work in a
comprehensive and reasonable way, and tackles theoretical as well as
practical aspects. There are two principal contributions of the paper:

1) A novel implementation of the Rate-Monotonic Scheduling algorithm,
which contains more complex and realistic details instead of the ideal
setting.

2) Rate-Monotonic Scheduling is investigated using Real-Time Maude for
the first time, to apply formal methods such as model checking and
theorem proving to analyze theoretical results, verify desired
properties, and evaluate results.

\ANSWER Thanks for the pros. We are happy that the work in the paper
interests the readers. We hope that it provides a new and formal way
to verify a real-time system, which is able to investigate important
problems (such as periodic task scheduling) from a more realistic
aspect.

\item
\COMMENT The weak point is the lack of evidence of practicality,
efficiency and scalability. Since there are only small and simplistic
scenarios to analyze this model, it is difficult to judge the
effectiveness of the proposed implementation. The key scenarios appear
conveniently small to allow the implementation fully supports all
these examples, returning at the same time appropriate counterexamples
to guide and adjust the design. Is this true for larger examples of
"real world" applications within the field of industrial real-time
systems? How do the authors propose to handle very large scenarios?
How does it behave on typical real-time scheduling problems used in
industry? I would like to see more complex examples to convince the
reader of the practicability of the approach on the assumptions given
for the model, and would probably increase the mentioned benefits. The
paper does not give any directions to the above problems, and some
sentences pointing out the difficulties in the implementation of more
generic and realistic scenarios would be most welcome to understand
the real contribution of the paper.

\ANSWER Thanks for the suggestion. As mentioned in the paper, the
system under verification is a real-world RMS implementation, which is
used in an avionic control system. The scenarios presented in Section
V.B are also real scenarios from the online system. Our industrial
partner has very strict requirements in the timing. They use at most
$5$ tasks for scheduling to avoid high overhead.

On the other hand, ``larger'' scenarios are not the same as
``complex'' scenarios in our model, as discussed in the revised
Section V.B. We have examined our approach by verifying randomly
generated scenarios where the number $n$ of tasks are over $10$. In
that case, the least common multiple of the periods of all tasks can
be as large as $10^5$ times the interrupt cycle $T$. All feasible
combinations of execution traces are checked in our approach. Our
industrial partner agrees that it satisfies their practical
requirements. Furthermore, the number $n$ of tasks can be $20$ and
even more if we allow different tasks possess a same period.

\hide{
the modeled RMS implementation serves in an avionic control systems
from our industrial partner. It is a real-world application within the
field of industrial real-time systems. And the scenarios presented in
Section V.B are also real scenarios that our partner is using in the
system. Our partner indicated that there would be at most $5$ tasks in
the task set for the implementation. On the other hand, in the current
version, we have randomly generated large numbers of test cases to
examine the efficiency of our schedulability test that is based on
model checking, as discussed in Section V.B. The capability of model
checking is restricted by the scale of the state space of the
model. In our model, the model checking can handle scenarios where
$mn$ is up to $10^6$ in an acceptable period of time. With $mn=10^6$
and reasonable task periods, the number $n$ of tasks can be at least
$10$, which has already fulfilled the requirements of the target
implementation. Further with particular task periods, the number $n$
of tasks can be $20$ or more.
}

\item
\COMMENT One of my biggest problems with the theoretical part of this
work is that the paper is technically solid as far as I can tell, but
there's not much yet in the way of theorems and results about formal
properties in Maude. The complete and detailed proof of the Theorem 2
should be provided somewhere. If the proof of the main results of
schedulability and correctness of the model is straightforward (as is
suggested) it seems not necessary to present a complex formalism on
Real-Time Maude and all the theorems provided are natural consequences
of Theorem 1. I agree that even non-surprising results need to be
investigated in the proof of Theorem 2, some sentences pointing out
the difficulties in the theoretical formalization in Real-Time Maude
would be most welcome. Please, include (e.g., in an appendix or even
in a technical report) a detailed proof.

\ANSWER We are happy that the readers are interested in the detailed
and technical proof, which is now included in the Appendix. It is true
that readers may be interested in the detailed proof of Theorem~2 to
feel convinced of the completeness of our approach. However, the proof
needs a bit more theoretical background about rewriting logic and more
technical details of our model, which we have simplified in the paper
in order to ease the understanding of our approach. We would prefer to
refer the detailed proof to a non-anonymous technical report, if
possible, in the final version.

\item
\COMMENT Another important point which needs improvement is the
description of the implementation. I guess other readers might have
less difficulties with the motivations if they know Maude and the
Real-Time Maude extension, but if you aim for a more general audience,
it might not work. Many pages are spent for the description of the
formal modeling of the implementation and the section is hard to
read. I highly recommend the authors to thoroughly revise this part of
the paper to make it easy-to-read. It would be better preferable to
have a much shorter description of the principal insights for the
implementation followed by more convincing examples and benchmarks
with respect to related work to convince the reader of the
practicability of the approach. You should try to summarize this
section and to extend the original contribution. From my point of
view, there is a trade off in this paper between the contributions and
the presentation of the implementation.

\ANSWER Thank you for the invaluable suggestions. Section IV has been
revised. Some unnecessary detailed definitions presented as code are
removed, and replaced by more explanations. We hope that the current
presentation would make Section IV more easy-to-read, even for an
audience who did not know Real-Time Maude before. On the other hand,
Section VI (Related Work) is also improved. We compare our work with
the existing theoretical work and the existing verification work in a
more complete way.

\end{itemize}

\subsection{Detailed Comments}
\begin{itemize}
\item
\COMMENT Page 7, Related Work. More comparisons with other approaches
based on model checking would help to evaluate pros and cons of an
implementation with Real-Time Maude with respect to different
languages and tools. The authors should review the comparison with
[24] and [25] more thoroughly. You should benchmark your approach
against others.

\ANSWER Thanks for the pointing this out. The related work section is
now improved. More comparisons with the theoretical approaches and
with [24,25] are added. The differences with [24,25] in the models and
in the way of modeling are mainly discussed. On the other hand, it is
a bit difficult to compare the efficiency and scalability between
[24,25] and our approach. One reason is that we have a different
setting with [24] and [25], because the objectives are different: we
aim at verifying an RMS implemention from a real-world real-time
system, while [24] and [25] targeted the RMS algorithm itself. Another
reason is that [24] and [25] presented no discussion on the efficiency
and scalability. Furthermore, they used \verb|TMSVL| and an extension
of SPIN, respectively, as verification tools, which are not
open-source, making us unable to re-implement their work.

\item
\COMMENT Page 8, Conclusions. Where in the paper do you show that the
details of your "realistic" implementation are "sufficient" for the
behaviors of all real systems used in the current industry? This
should be briefly discussed.

\ANSWER Sorry for that. The statement has been corrected. The work
presented in the paper aims at modeling and verifying a realistic RMS
implementation in an industrial avionic system. It is a piece of real
work in the industry. Even though we believe that our approach could
be applied to other similar systems as well, it means more to that
specific one. In this sense, ``sufficient'' means that the details
described in our model meet the requirements and expectations from the
users and the engineers.

\end{itemize}

\section{Responses to Reviewer 2}
\subsection{General Comments}
\begin{itemize}
\item
\COMMENT One of the problems with this paper is that it is not
completely self-contained.  Some knowledge about Real-Time Maude is
necessary in order to fully understand it.  I found the short
introduction about Real-Time Maude (section II.B) not enough for
understanding the formalism used in the rest of the paper (see details
below).

\ANSWER Thank you very much for pointing this out. The Sections II.B
and IV are improved to be more self-contained now. The idea is that,
Section II.B presents mainly an overview of Real-Time Maude, and then
Section IV describes the model with introducing necessary syntax
on-the-fly.

\item
\COMMENT Another concern is that the significance of the paper could
be much better if the paper addressed the problem of modelling RMS in
a more general way. Instead, the paper takes the form of a case study,
where the formal modelling of one specific RMS implementation is
presented, rather than a general method for modelling and analyzing
RMS implementations.

\ANSWER Thank you for the suggestion. Yes, we agree that we have two
choices to do the work presented in the paper: one is to develop a
general way to model RMS algorithms or implementations, and then
instantiate the general model to get a concrete one for our target
system; the other is to shape an accurate model of the target system
directly.

To us, there is a trade off between these two directions. We did
choose the latter according to the following considerations. By our
experience, \emph{the development engineers (who write the code) and
  verification engineers (who verify the code or the system) are
  usually not the same persons.} When a verification engineer models a
system, he is actually abstacting the system or the codes. It is
difficult to ensure that the model behaves the same as the system or
the codes.  And it is more difficult to make other engineers and users
believe that the model behaves the same as the system or the
codes. Furthermore, the developers and users hope that the model
behaves not only the same as what they had in mind, but also the same
as what is written in the code. A way out is to make the model not
only \emph{behave like} the code, but also \emph{``look'' like} the
code. This requires the model to be as accurate as possible, based on
a fact that the expressivity of the modeling language is powerful
enough. It makes the model specific, but improves the engineers'
confidence in the model and the verification results. That is also why
we emphasize the correspondences between the functions in the model
and the lines in the pseudocode, when we introduce our model in
Section~IV.

On the other hand, we are also interested in the first direction. This
will be the future work and we would try to get a better balance
between them.

\item
\COMMENT Also, the discussion about soundness and completeness of the
analysis (section V.C) is not fully developed and, once again, it
refers to the specific model rather than being a general result.

\ANSWER Thanks for this comment. The section about soundness and
completeness is now made more self-contained. The detailed proof of
Theorem~2 is included in the Appendix. On the other hand, since we
chose the accurate way to achieve the model as discussed above, yes,
the proof is a specific one.

\end{itemize}

\subsection{Detailed Comments}
\begin{itemize}
\item
\COMMENT [Section II.B] In the definition of $IR$, it is not clear
what $s$ and $s'$ are. I guess they are states, but the concept of
state was not introduced.

\ANSWER Yes, they can be states. But more generally, they are
terms. This part has been modified to specify $s$ and $s'$ (in fact,
$t$ and $t'$ now) clearly.

\item 
\COMMENT [Section II.B] In the definition of a class, the authors
should specify that a class includes a collection of rules, otherwise
this fact may be missed.

\ANSWER Thanks very much for reminding. This is added now.

\item
\COMMENT [Section III] Page 2 right column, line -24: The
implementation is shown as schedule() in Figure 1 $\rightarrow$ The
pseudocode of schedule() is shown in Figure 1

\ANSWER Thanks for the careful reading. It is done now.

\item 
\COMMENT [Section IV.A] I don't understand the meaning of
\verb|[ctor]| in the definition of some operations (maybe because I
have no specific background on Maude). I could not find the
explanation of this writing in the introduction to Maude.

\ANSWER Sorry for missing the explanation of \verb|[ctor]|. It is now
added when \verb|[ctor]| is used for the first time.

\item
\COMMENT [Section IV.A] For the stack sort, the authors mention
operations push, pop and peek, but then you don't write the
definitions of these operations. Instead, they write the definitions
of bottom and \#, which are not explained in the text. This part
should be made clearer.

\ANSWER Sorry for the misleading in the text. In fact, \verb|bottom|
and \verb|#| are constructors of sort \verb|Stack|. This part is
modified now. Some detailed definitions in code are removed, and are
replaced by more explanations. We hope that the current presentation
would make Section IV more easy-to-read.

\item
\COMMENT [Section IV.A] Also, the meaning of NzNat is not explained.

\ANSWER Sorry for this incompleteness. \verb|NzNat| did mean non-zero
natural numbers. \verb|NzNat| is now replaced by \verb|Nat| to reduce
unnecessary complexity for the readers.

\item
\COMMENT [Section V.C] In the statement of Theorem~1, the definition
of time-robust real-time rewrite theory is missing. Also, the meaning
of tick-stabilizing atomic propositions is not defined. This makes the
section not self-contained.

\ANSWER Sorry for the inconvenience for reading. In fact, the
definitions of time-robustness and tick-stabilization (which is now
replaced with tick-invariance) require more knowledge about rewriting
logic. We avoid introducing the accurate definitions of them. Instead,
now we give short descriptions of them before Theorem~1. More detailed
introduction can be found in the Appendix.

\item
\COMMENT [Section V.C] The proof of theorem 2 is missing. The authors
just indicate how the proof could be developed, but they do not
develop it. A possibility would be to point to a document where this
proof has been developed.

\ANSWER We are very happy that the readers are interested in the
detailed and technical proof, which is now included in the Appendix.
The detailed proof requires a bit more theoretical background about
rewriting logic and more details of our model, which we tried to
simplify in the paper in order to ease the understanding of our
approach. We would prefer to refer the detailed proof to a
non-anonymous technical report, if possible, in the final version.
\end{itemize}

\section{Responses to Reviewer 3}
\begin{itemize}
\item
\COMMENT RMS has many different realization methods, no comparison is
given to justify the strong point of real-time maude when applied to
RMS.

\ANSWER The comparison could be found in Section VI (and we have
expanded according to the reviewers' comments :-)). We would like to
clarify that we are \emph{not implementing} RMS, we are
\emph{verifying} it. We choose Real-Time Maude since it verifies the
model using model checking technique. If verification passes, no
deadline will be missed in the implementation as long as the model
assumptions are met.

\hide{
The comparison is given in the related work section (Section
VI). We would say that we are not implementing RMS algorithm in
Real-Time Maude. In fact, we are using Real-Time Maude to model an
realistic implementation of RMS from an industrial avionic
system. There is no theoretical reason telling us that we have to use
Real-Time Maude to model the RMS implementation. However, the existing
verification work as far as we know cannot fit our needs, as discussed
in Section VI. Therefore, we developed the work in the paper, as a
piece of real work in the industry.
}

\item
\COMMENT Rate-monotonic scheduling is a very simple scheduling and has
been sufficiently investigated in the field of embedded system, the
authors claim the innovation of this paper as the implementation using
real-time maude, however, I cannot see any special point (or
advantage) when using this real-time maude in modelling RMS.

\ANSWER Thank you for the insightful comment. We agree that RMS has
been intensively studied, because of its importance. However, most of
the results are based on the assumption that \emph{task switching does
  not take time}. That assumption is not realistic. Models with and
without the assumption have totally different behaviors. For instance,
as shown in Figure~2(b), in the RMS implementation we verified, the
switching from time $9$ to $11$ blocks the interrupt handling, hence
delaying the initiation time of $\tau_1$. To our knowledge, this kind
of phenomenon would not happen in the theoretical analysis
models. Although there is very few results that take into account the
switching overhead (as discussed in Section VI), none of them is
generic and our system under verification is not in their scope.

Despite the above reason, the most important motivation of this paper
is to verify the system not only for the schedulability, but also for
the correctness. That is, we want to verify that the target system
does schedule the tasks under the RMS algorithm. The work presented
should be seen as a verification problem, instead of a schedulability
test. From verification point of view, this paper models a real-world
RMS implementation with sufficient technical details. Two important
properties--schedulability and correctness--are verified by model
checking technique, and the soundness and completeness of the results
are demonstrated.

\hide{
is a very simple scheduling and is the optimum at the same time among
the fixed-prority ones, making it widely applied and sufficiently
investigated. Besides the ideal setting, the schedulability of RMS was
also investigated under different kinds of implementations with
consideration about overhead, as we discuss in the related work
section (Section VI). Despite the reason that the existing results do
not fit our target system using RMS, the most important motivation to
develop the work presented is that, we aim at verifying the system not
only for the schedulability, but also for the correctness. That is, we
want to verify that the target system does schedule the tasks under
the RMS algorithm. The work presented should be seen as a verification
problem, instead of a schedulability test.
}

\item
\COMMENT The assumptions given in IV is too rigid, which makes the
scheduling problem to be investigated by very simple, then what is the
difficulty?

\ANSWER Thank you for the question. When an automatic formal
verification tool (such as model checker) is used to verify a given
system, it explores the whole system state space to check whether the
given property holds. Then the state space explosion is the most
common problem for an automatic verification tool. That means the
state space of the system/model is too huge for the tool to explore
with acceptable temporal and spatial cost. The scale of the state
space depends on the complexity of the model of the system.

As an industrial verification application, we are not trying to make
the problem as difficult as possible. Instead, we are trying to
simplify the model so that the existing tools are able to solve, while
the target system satisfies the assumptions on the model so that the
verification results are trustworthy.

On the other hand, the assumptions are not so rigid. For example, the
mask mechanism considered makes the job initiation times of task
$\tau_i$ possibly not equal $kT_i$, implying that the deadlines of the
jobs may be not equal to $(k+1)T_i$ (see assumption A1'). This is also
discussed in the beginning of Section IV, pointing out the blocking at
time $10$ in Figure~2(b). Another example is that non-determinism
exists in our model, as discussed in Section~VI.

\item
\COMMENT The description in section IV (A-E) is trivial, it can only
be viewed as a simple application of the rewriting logic.

\ANSWER Thanks for the careful reading. We agree that this is an
application of rewriting logic. In this paper, we do apply the
existing theories and techniques of rewriting logic to solve a real
industrial problem, which has not yet been investigated by this sort
of techniques, achieving an industrial contribution.

\item
\COMMENT The model checking part is not sufficiently discussed, the
authors only list the commands but not illustrate the mechanism behind
these commands.

\ANSWER Thank you for the suggestion. However, since the model checker
is not part of our contributions in the paper, we do not think it is
appropriate to illustrate the mechanism behind the model checking
commands, which may confuse the readers on the other hand.

\end{itemize}

\hide{
% use section* for acknowledgment
\ifCLASSOPTIONcompsoc
  % The Computer Society usually uses the plural form
  \section*{Acknowledgments}
\else
  % regular IEEE prefers the singular form
  \section*{Acknowledgment}
\fi

The authors would like to thank...
}

% Can use something like this to put references on a page
% by themselves when using endfloat and the captionsoff option.
\ifCLASSOPTIONcaptionsoff
  \newpage
\fi

% trigger a \newpage just before the given reference
% number - used to balance the columns on the last page
% adjust value as needed - may need to be readjusted if
% the document is modified later
%\IEEEtriggeratref{8}
% The "triggered" command can be changed if desired:
%\IEEEtriggercmd{\enlargethispage{-5in}}

% references section

% can use a bibliography generated by BibTeX as a .bbl file
% BibTeX documentation can be easily obtained at:
% http://www.ctan.org/tex-archive/biblio/bibtex/contrib/doc/
% The IEEEtran BibTeX style support page is at:
% http://www.michaelshell.org/tex/ieeetran/bibtex/
%\bibliographystyle{IEEEtran}
% argument is your BibTeX string definitions and bibliography database(s)
%\bibliography{IEEEabrv,submission}
%
% <OR> manually copy in the resultant .bbl file
% set second argument of \begin to the number of references
% (used to reserve space for the reference number labels box)

% biography section
% 
% If you have an EPS/PDF photo (graphicx package needed) extra braces are
% needed around the contents of the optional argument to biography to prevent
% the LaTeX parser from getting confused when it sees the complicated
% \includegraphics command within an optional argument. (You could create
% your own custom macro containing the \includegraphics command to make things
% simpler here.)
%\begin{IEEEbiography}[{\includegraphics[width=1in,height=1.25in,clip,keepaspectratio]{mshell}}]{Michael Shell}
% or if you just want to reserve a space for a photo:

\hide{

\begin{IEEEbiography}{Jiaxiang Liu}
Biography text here.
\end{IEEEbiography}

% if you will not have a photo at all:
\begin{IEEEbiographynophoto}{John Doe}
Biography text here.
\end{IEEEbiographynophoto}

% insert where needed to balance the two columns on the last page with
% biographies
%\newpage

\begin{IEEEbiographynophoto}{Jane Doe}
Biography text here.
\end{IEEEbiographynophoto}
}

% You can push biographies down or up by placing
% a \vfill before or after them. The appropriate
% use of \vfill depends on what kind of text is
% on the last page and whether or not the columns
% are being equalized.

%\vfill

% Can be used to pull up biographies so that the bottom of the last one
% is flush with the other column.
%\enlargethispage{-5in}

% that's all folks
\end{document}