### 2016-11-09 RMS/final-version/TXT_15-TIE-3480.tex
included bbl into TXT.tex

%\documentclass[onecolumn]{IEEEtranTIE}
\documentclass[journal]{IEEEtranTIE}
\usepackage{graphicx}
\usepackage[noadjust]{cite}
\usepackage{picinpar}
\usepackage{amsmath}
\usepackage{stfloats}
\usepackage{url}
\usepackage{flushend}
\usepackage[latin1]{inputenc}
\usepackage{colortbl}
\usepackage{soul}
\usepackage{multirow}
\usepackage{pifont}
\usepackage{color}
\usepackage{alltt}
\usepackage[hidelinks,bookmarks=false]{hyperref}
\usepackage{enumerate}
\usepackage{siunitx}
\usepackage{breakurl}
\usepackage{epstopdf}
\usepackage{pbox}

\newtheorem{theorem}{Theorem}
\newtheorem{lemma}[theorem]{Lemma}
\newtheorem{definition}[theorem]{Definition}

% Define the fontsize in environment {verbatim}
\makeatletter
\def\verbatim{\small\@verbatim \frenchspacing\@vobeyspaces \@xverbatim}
\makeatother

\usepackage{microtype}

\begin{document}
\title{Formal Modeling and Verification of a Rate-Monotonic Scheduling Implementation with Real-Time Maude}

\author{
  \vskip 1em
  Jiaxiang~Liu,
  Min~Zhou,
  Xiaoyu~Song,
  Ming~Gu,
  and Jiaguang~Sun

  \thanks{ 
    Manuscript received November 8, 2015; revised March 30,
    2016 and July 13, 2016; accepted October 20, 2016. This work was
    supported in part by NSFC Program (No.\ 91218302, No.\ 61527812,
    No.\ 61402248), National Science and Technology Major Project
    (No.\ 2016ZX01038101), MIIT IT funds (Research and application of
    TCN key technologies) of China, and The National Key Technology
    R\&D Program (No.\ 2015BAG14B01-02).

    J.\ Liu is with the Department of Computer Science \& Technology,
    Tsinghua University, Beijing, China (e-mail:
    jiaxiang.liu@hotmail.com).

    M.\ Zhou is with the School of Software, Tsinghua University,
    Beijing 100084, China (corresponding author; phone:
    +86-10-6277-3275; fax: +86-10-6278-3549; e-mail:
    zhoumin03@gmail.com).

    X.\ Song is with the Department of Electrical \& Computer
    Engineering, Portland State University, Portland, OR 97207-0751,
    USA.

    M.\ Gu and J.\ Sun are with the School of Software, Tsinghua
    University, Beijing 100084, China.

    J.\ Liu, M.\ Zhou, M.\ Gu and J.\ Sun are also with Key Laboratory
    for Information System Security, Ministry of Education, China, and
    Tsinghua National Laboratory for Information Science and
    Technology, China.  
  }  
}

\maketitle

\begin{abstract}
Rate-monotonic scheduling (RMS) is one of the most important real-time
scheduling used in industry. There are a large number of results about
RMS, especially on its schedulability. However, the theoretical
results do not contain enough details to be used directly for an
industrial RMS implementation. On the other hand, the correctness of
such an implementation is of the crucial importance. In this paper, we
analyze a realistic RMS implementation by using Real-Time Maude, a
formal modeling language and analysis tool based on rewriting
logic. Overhead and some details of the hardware are taken into
account in the model. We validate the schedulability and the
correctness of the implementation within key scenarios. The soundness
and the completeness of our approach are substantiated.
\end{abstract}

\begin{IEEEkeywords}
Embedded systems, formal verification, modeling, real-time systems,
rewriting logic, scheduling.
\end{IEEEkeywords}

\markboth{IEEE TRANSACTIONS ON INDUSTRIAL ELECTRONICS}%
{}

\definecolor{limegreen}{rgb}{0.2, 0.8, 0.2}
\definecolor{forestgreen}{rgb}{0.13, 0.55, 0.13}
\definecolor{greenhtml}{rgb}{0.0, 0.5, 0.0}

\section{Introduction}\label{s:introduction}

\IEEEPARstart{P}{eriodic} task scheduling is one of the most important
topics within the field of industrial real-time systems. A set of
periodic tasks is said to be \emph{schedulable} with respect to some
scheduling algorithm if all jobs meet their
deadlines. \emph{Rate-monotonic scheduling} (\emph{RMS}) is a
\emph{fixed} priority scheduling algorithm for preemptive hard
real-time environments proposed by Liu and
Layland~\cite{DBLP:journals/jacm/LiuL73}, which assigns priorities to
jobs according to the periods of the corresponding tasks: the smaller
period, the higher priority. RMS is proven to be the \emph{optimal}
fixed priority scheduling algorithm~\cite{DBLP:journals/jacm/LiuL73},
in the sense that any set of tasks, which is schedulable under
\emph{some} fixed priority scheduling algorithm, is also schedulable
with respect to RMS. It is widely used in safety-critical real-time
applications, such as vehicles and avionics, thanks to its optimality
and easiness to implement.

Liu and Layland~\cite{DBLP:journals/jacm/LiuL73} gave a sufficient
condition for the schedulability of a set of $n$ tasks scheduled by
RMS: $\Sigma^n_{i=1}C_i/T_i \le n(2^{1/n}-1)$, where $C_i$ and $T_i$
are the \emph{computation (time) requirement} and the period of task
$\tau_i$, respectively. Two main directions on RMS have been explored
since then. One is to relax the assumptions on the original RMS model,
making it applicable on more systems.  For instance,
\cite{DBLP:conf/rtss/LehoczkySS87,DBLP:journals/rts/SpruntSL89,DBLP:conf/rtss/LehoczkyR92,DBLP:journals/tc/StrosniderLS95}
allow aperiodic tasks in the scheduling,
\cite{DBLP:journals/pe/LeungW82,audsley1993deadline} generalize RMS to
be \emph{deadline-monotonic}, \cite{DBLP:journals/tc/ShaRL90} allows
resource sharing among tasks,
\cite{dhall1978real,DBLP:journals/rts/LopezGDG03,DBLP:journals/tpds/LopezDG04,DBLP:journals/tc/BaruahG03}
extend RMS on multiprocessors, and
\cite{DBLP:journals/rts/OhS94,DBLP:journals/rts/GhoshMMS98,DBLP:journals/tpds/BertossiMR99}
enhance fault-tolerance. The other direction is to generate better
schedulability test conditions for the algorithm and its
extensions~\cite{DBLP:conf/rtss/LehoczkySD89,DBLP:conf/rtss/KuoM91,DBLP:journals/tc/BiniBB03,DBLP:journals/rts/LopezGDG03,DBLP:journals/tc/BaruahG03,gardner1999}. The
RMS algorithm is no doubt of practical importance.

It is even more crucial to ensure the reliability of an implementation
instead of the algorithm, when RMS serves in a safety-critical
system. When analyzing a realistic implementation, theoretical results
may be no more applicable. For instance, even though the conditions
derived from algorithm analysis are satisfied, schedulability can be
broken by overhead in the system, or by the interrupt mask mechanism
which may delay interrupt handling. On the other hand, correctness of
the implementation with respect to the algorithm is difficult to be
verified by the traditional methods such as testing and simulation due
to their incompleteness. Extensive effort to apply formal methods,
such as model checking and theorem proving, has been made to analyze
safety-critical systems for the past few
years~\cite{DBLP:journals/tie/JiangZLDSGS15,DBLP:journals/iandc/MeseguerR13,DBLP:journals/cacm/Leroy09,DBLP:conf/sosp/KleinEHACDEEKNSTW09}. However,
as far as we know, few~\cite{TianD2011,DBLP:conf/iceccs/CuiDT14}
attempt to analyze the RMS algorithm, while no work for
implementations of RMS is found.

In this paper, we use \emph{Real-Time Maude}, a \emph{rewriting}-based
formal modeling language and analysis tool for real-time systems, to
model a realistic implementation of RMS that serves as a simplified
operating system within an avionic control system, and then verify
several desired properties. Based on a realistic implementation, our
model extends the standard RMS model proposed
in~\cite{DBLP:journals/jacm/LiuL73}, by taking into account overhead
and other details of the hardware platform.

The rest of this paper is organized as
follows. Section~\ref{s:background} gives some background of both the
RMS algorithm and Real-Time Maude.  Section~\ref{s:imp} presents the
RMS implementation that we model and analyze.
Section~\ref{s:formalism} introduces how we model the RMS
implementation using Real-Time Maude. Then
Section~\ref{s:verification} explains how to verify the desired
properties and to evaluate the results. Related work is discussed in
Section~\ref{s:relate}. We conclude the paper in
Section~\ref{s:conclusion}.

\section{Background}
\label{s:background}

\subsection{Rate-Monotonic Scheduling Algorithm}
\label{ss:rms}
A task set consists of \emph{only} $n$ periodic tasks
$\tau_1,\ldots,\tau_n$. Each task $\tau_i$ has a period $T_i$ and a
computation requirement $C_i$. First jobs of all tasks are assumed to
be initiated at time $0$ simultaneously.  Deadlines consist of
runnability constraints only: the deadline of a job corresponding to
$\tau_i$ is the initiation time of the next job corresponding to
$\tau_i$.  The RMS algorithm chooses the labeling such that $T_1\le
T_2\le \ldots \le T_n$. Consequently, $\tau_i$ is given priority $i$,
assuming smaller numbers have higher priorities. The following
assumptions are made:

(A1) Jobs corresponding to task $\tau_i$ are initiated exactly at
times $kT_i$ with integers $k\ge 0$.

(A2) Computation requirement $C_i$ for each task $\tau_i$ is constant
and does not vary with time.

(A3) Tasks are independent, such that they are ready to run at their
initiation times and can be preempted instantly (ignoring all
blocking).

(A4) All overhead, such as task switching time, is ignored.

A simple example showing this algorithm is depicted in
Fig.~\ref{f:example}(a). However in this paper, we consider an
implementation instead of the RMS algorithm itself, thus the model
would be more complicated than this standard, ideal setting.  (A1)
will be modified because of the interrupt mask mechanism, while (A4)
is relaxed to obtain a more realistic analysis model.

\subsection{Real-Time Maude}
Real-Time Maude~\cite{DBLP:journals/lisp/OlveczkyM07} is an extension
of \emph{Maude}~\cite{DBLP:journals/tcs/ClavelDELMMQ02} which is a
language and tool based on \emph{rewriting
  logic}~\cite{DBLP:journals/jlp/Meseguer12}.  It supports formal
specification and analysis of real-time systems.

\subsubsection{Specification}
Real-Time Maude models systems as \emph{modules}. A module specifies a
\emph{real-time rewrite theory} $\mathcal{R} = (\Sigma,
\mathit{E}\cup\mathit{A} , \mathit{IR}, \mathit{TR})$, where:
\begin{itemize}
\item $\Sigma$ is an algebraic \emph{signature}, that is, a set of
  declarations of \emph{sorts}, \emph{subsorts} and \emph{function
    symbols}. The function symbols are allowed to be mixfix, in which
  case underscores ``\verb|_|'' indicate the positions of parameters.
  \emph{Terms} are expressions built from function symbols and
  variables.
\item $(\Sigma, \mathit{E}\cup\mathit{A})$ is a \emph{membership
  equational logic theory}, with $\mathit{E}$ a set of
  \emph{conditional equations} and \emph{memberships} on $\Sigma$, and
  $\mathit{A}$ a set of equational axioms such as associativity,
  commutativity and identity.  $(\Sigma, \mathit{E}\cup\mathit{A})$
  models the system's ``static'' states as terms of some sort, and is
  equipped with a built-in specification of a sort \verb|Time|.
\item $\mathit{IR}$ is a set of \emph{labeled conditional rewrite
  rules} specifying the system's local transitions. Each rule has the
  form $[l]~:~t\rightarrow t'\mbox{ \textbf{if}
  }\bigwedge^n_{j=1}cond_j$, where each $cond_j$ is an equality
  $u_j=v_j$, and $l$ is a \emph{label}, $t,t',u_j,v_j$ are terms. Such
  a rule specifies an \emph{instantaneous transition}, without
  consuming time, from an instance of $t$ to the corresponding
  instance of $t'$, \emph{provided} the conditions hold.
\item $\mathit{TR}$ is a set of \emph{(labeled) tick rules} of
  the form $[l]~:~\{s\}\rightarrow\{s'\} \mbox{\textbf{ in time}
  }r\mbox{ \textbf{if} }cond$ that specify \emph{timed
    transitions}. Each tick rule advances time by $r$ time units from
  the \emph{entire} state modeled by term $s$ to the destination state
  $s'$.
\end{itemize}
$\mathit{IR}$ and $\mathit{TR}$ together model the system's ``dynamic''
behaviors.

In rewriting logic, rewrite rules are applied non-deterministically,
that is, when several rules can be applied on a given term $t$, any of
them may be chosen. Hence non-deterministic behaviors can be modeled
naturally in Real-Time Maude.  Real-Time Maude also supports
specifications in \emph{object-oriented} style.  A class declaration
$\texttt{class }C\texttt{ |
}att_1\texttt{:}s_1\texttt{,}\ldots\texttt{,}att_n\texttt{:}s_n$
defines a class $C$ with attributes $att_1$ to $att_n$ of sorts $s_1$
to $s_n$, respectively. An \emph{object} of class $C$ is represented
as a term $\texttt{< } O\texttt{:} C \texttt{ | }
att_1\texttt{:}val_1\texttt{,} \ldots
\texttt{,}att_n\texttt{:}val_n\texttt{ >}$ of sort \verb|Object|,
where $O$, of sort \verb|Oid|, is the object's \emph{identifier}, and
$val_i$ is the value of the attribute $att_i$ with $i\in [1,n]$. Rules
can be defined on a given class. A \emph{subclass} inherits all the
attributes and rules of its superclasses.

\subsubsection{Formal Analysis}
Real-Time Maude provides many useful commands and tools to analyze a
given model. For example, \verb|rewrite| allows to execute the model,
symbolically; given an initial state, \verb|search| is used to search
reachable states satisfying desired properties; the Maude's
\emph{Inductive Theorem Prover} (ITP) can be applied to interactively
prove properties written in \emph{membership equational logic}.

In this paper, we only consider Real-Time Maude's \emph{Linear
  Temporal Logic (LTL) model checker}, which analyzes whether
\emph{each} behavior satisfies a temporal logic formula. \emph{State
  propositions} are defined as terms of sort \verb|Prop|. Their
semantics is defined by conditional equations of the form $\texttt{ceq
} \mathit{statePattern} \texttt{ |= } \mathit{prop} \texttt{ = } b
\texttt{ if } cond$, with $b$ a term of sort \verb|Bool|, stating that
$\mathit{prop}$ evaluates to $b$ in states which are instances of
$\mathit{statePattern}$ provided the condition $cond$ holds. These
equations together define $\mathit{prop}$ to hold in all states $s$
that make $s \texttt{ |= } \mathit{prop}$ evaluate to \verb|true|. A
temporal logic \emph{formula} is constructed by state propositions and
temporal logic operators such as \verb|~|(negation),
\verb|\/|(disjunction), \verb|[]|(``always''),
\verb|U|(``until''). Real-Time Maude supports both \emph{timed} and
\emph{untimed LTL model checking}. The untimed model checking command
\begin{alltt}
  (mc \(s\) |=u \(\Phi\) .)
\end{alltt}
checks whether the temporal logic formula $\Phi$ holds in all
behaviors starting from the initial state $s$, \emph{with no time
  limit}.

\section{The Implementation of RMS}
\label{s:imp}
The implementation written in C is from an industrial avionic control
system.  Interrupts would be triggered by, and only by, the clock
every $T$, which we call \emph{interrupt cycle}. When an interrupt
request occurs, if the system is interruptible, i.e. the interrupt
mask bit is cleared, the handler function $\mathit{schedule()}$ will
be invoked; otherwise $\mathit{schedule()}$ will be pending until the
interrupt mask bit becomes cleared.  The pseudocode of
$\mathit{schedule()}$ is shown in Fig.~\ref{f:schedule}, where
$\mathit{taskList}$ is the list of periodic tasks to be scheduled. We
assume that the list is in descending order of priority, and both
variables $\mathit{taskList}$ and $\mathit{timer}$ are global. In this
implementation, there is only one kind of interrupt, the period $T_i$
of each task is a multiple of $T$, and the tasks are independent,
meeting assumption~(A3).

\begin{figure}[!t]
\centering
\includegraphics[width=8.8cm]{FIG1_15-TIE-3480.pdf}
\caption{The C-Like Pseudocode of $schedule()$}
\label{f:schedule}
\end{figure}

In Fig.~\ref{f:schedule}, the handler function $\mathit{schedule()}$
first updates status of all tasks in $\mathit{taskList}$ via function
$\mathit{updateStatus()}$. This updating actually initiates tasks that
should be scheduled in the current interrupt cycle. Then
$\mathit{schedule()}$ traverses the list to execute the ready tasks
one by one, or to do a return when encountering an
interrupted\footnote{Note that the status \textit{INTERRUPT} indicates
  the task is interrupted currently, or was interrupted before but its
  execution is not complete yet.}  task. As for the function
$\mathit{updateStatus()}$, it updates each task in two steps: firstly,
if the task is running, it becomes interrupted; secondly for the task
at its initiation time, if its previous job is complete, it would be
set ready, otherwise it misses its deadline, producing an
error. Notice that $\mathit{schedule()}$ is invoked only when the
interrupt request is handled, not when the interrupt is disabled (the
interrupt mask bit is set). Due to the interrupt mask bit, its
execution cannot be interrupted when it is updating status of tasks or
searching the next task to execute, however, it can be interrupted
while executing some task (Line~12). This allows the execution of
$\mathit{schedule()}$ to be nested.

For simplicity, in the rest of this paper, we use \emph{scheduling} to
refer to the stage, from the moment when a pending interrupt request
is detected, to the moment when the first should-be-run periodic task
starts executing, i.e. Line~8 or~12 in
Fig.~\ref{f:schedule}. Therefore, \emph{scheduling time} mainly
consists of three parts: (i)~the time for switching context from the
running task, possibly none, to $\mathit{schedule()}$ when an
interrupt request is handled, (ii)~the time spent by
$\mathit{schedule()}$ searching and setting the \emph{first}
should-be-run periodic task (Lines~2--11 in Fig.~\ref{f:schedule}),
and (iii)~the time for switching context from $\mathit{schedule()}$ to
that task. \emph{Switching} refers to the stage, from the moment when
a periodic task completes its execution, to the moment when the next
should-be-run periodic task starts executing. \emph{Switching time}
thus also consists of three parts: (i)~the time for switching context
from the complete task back to $\mathit{schedule()}$, (ii)~the time
spent by $\mathit{schedule()}$ searching and setting the \emph{next}
should-be-run periodic task, and (iii)~the time for switching context
from $\mathit{schedule()}$ to that task.

\section{Formal Modeling of the Implementation}
\label{s:formalism}
Considering some technical details of the platform, such as interrupt
mask mechanism, we model the implementation under the following
assumptions:

(A1') Jobs corresponding to task $\tau_i$ are initiated at the
beginning of scheduling that handles the requests triggered at times
$kT_i$ with integers $k\ge 0$.

(A2) Computation requirement $C_i$ for each task $\tau_i$ is constant
and does not vary with time.

(A3) Tasks are independent, such that they are ready to run at their
initiation times and can be preempted instantly.

(A4') Scheduling time and switching time are considered, while other
overhead is ignored.

These assumptions make our model different from the standard one. For
instance, with~(A1'), an interrupt request that occurs during
switching will be pending, such that jobs of task $\tau_i$ cannot
initiate at $kT_i$. They should wait until switching finishes and the
interrupt mask bit is cleared, which is different from~(A1). (A3) says
that jobs are ready to run at their initiation times, however, no one
can start running at exactly its initiation time, because scheduling
takes time. Under these assumptions, an example showing the execution
of tasks scheduled by the implementation is depicted in
Fig.~\ref{f:example}(b). Note that the second job instance of $\tau_1$
is initiated at time $11$ instead of $10$, because switching is
performed and the interrupt mask bit is set at time $10$, exemplifying
differences between (A1') and (A1).

\begin{figure}[!t]
\centering
\includegraphics{FIG2_15-TIE-3480.pdf}
\caption{A set of $2$ periodic tasks scheduled under (a) RMS algorithm
  and (b) the RMS implementation, respectively. $T=T_1=10$, $T_2=20$,
  $C_1=3$, $C_2=2$. In~(b), both scheduling time and switching time
  are $2$.}
\label{f:example}
\end{figure}

In this section, we introduce first how we model states -- the static
aspect -- of the system using terms of given sorts, or called data
types, then how we specify essential behaviors -- the dynamic aspect
-- using rewrite rules. In particular, modeling of \emph{instantaneous
  behaviors} would be explained in Sections~\ref{ss:ir}, \ref{ss:init}
and~\ref{ss:inthandling}, followed by \emph{timed behaviors} in
Section~\ref{ss:timedbehavior}.

\subsection{Basic Data Types}
In our model, tasks are identified by their indexes of sort \verb|Nat|
in $\mathit{taskList}$. We define a sort \verb|MaybeNat| wrapping
\verb|Nat|s to refer to some task, with \emph{constructor} \verb|some|
followed by a \verb|Nat| $n$ indicating the task indexed $n$, and
\verb|none| for no task:
\begin{verbatim}
  op none : -> MaybeNat [ctor] .
  op some_ : Nat -> MaybeNat [ctor] .
\end{verbatim}
where the keyword \verb|ctor| denotes the corresponding function
symbol to be a constructor.

A sort \verb|Stack| is introduced to model the stack of the system,
storing the tasks that are being interrupted, and equipped with
operations \verb|push|, \verb|pop| and \verb|peek| on it.

We also need a sort \verb|Counter| to record the execution of tasks.
We call \emph{execution time} the time how long a task has been
executing for. A \verb|Counter| records the execution time and the
computation requirement of a task.

The global variable $\mathit{timer}$ is reset when it reaches an upper
bound while increasing, which is not shown in detail in
Fig.~\ref{f:schedule} but is reasonable. The upper bound is the least
common multiple of periods of all tasks. A sort \verb|Timer| is
defined to model $\mathit{timer}$.

\subsection{Modeling the System States}
The system can be considered as consisting of several parts: the tasks
which are scheduled, the scheduler itself, the hardware including
registers and stacks, and the interrupt source. The scheduler,
i.e.\ the function $\mathit{schedule()}$, can be described by a single
variable $\mathit{timer}$. We present the models of the other parts
one by one.

\subsubsection{Tasks}
Each task is abstracted from its functionality as a \verb|Counter|.
Overhead for scheduling and switching is considered in our model. They
are treated as two system tasks. Every task is modeled as an object
instance of some subclass of the base class \verb|Task|:
\begin{verbatim}
  class Task | cnt : Counter .
  op error : -> Object [ctor] .
\end{verbatim}
where \verb|error| is an object indicating some task that misses its
deadline.

A periodic task, which needs to be scheduled, is an object instance of
the subclass \verb|PTask| of \verb|Task| with additional attributes
\verb|priority|, \verb|period| and \verb|status|:
\begin{verbatim}
  class PTask | priority : Nat, period : Nat, 
                status : Status .
  subclass PTask < Task .
\end{verbatim}
where \verb|Status| is a sort with four constant constructors
\verb|RUNNING|, \verb|INTERRUPT|, \verb|READY| and \verb|DORMANT|,
same as in the implementation.

The list of periodic tasks, the variable $\mathit{taskList}$ in the
implementation, is modeled as an instance of sort \verb|TaskList|,
which is a list of \verb|PTask|s and/or \verb|error|s.  Periodic tasks
are identified by their indexes in the list.

On the other hand, a system task is an object instance of the subclass
\verb|SysTask| of \verb|Task| with no extra attributes.  Different
from periodic tasks, system tasks are organized in a multiset of sort
\verb|SysTasks|, and identified by their \verb|Oid|s.

\subsubsection{Hardware}
Our model considers two parts of hardware related to interrupt
handling: the registers and the stack.

The set of registers is modeled as an object instance of class
\verb|Regs|, with attributes \verb|pc| denoting the program counter,
\verb|mask| for the interrupt mask bit and \verb|ir| for the interrupt
request bit:
\begin{verbatim}
  class Regs | pc : TaskID, 
               mask : Bool, ir : Bool .
\end{verbatim}
where the sort \verb|TaskID| contains subsorts \verb|MaybeNat| and
\verb|Oid|, referring to some task. Some operations, such as
\verb|getPc| and \verb|setMask|, are defined on the class \verb|Regs|.

Then the hardware is described by the sort \verb|Hardware| consisting
of an instance of \verb|Regs| and a term of sort \verb|Stack|.

\subsubsection{Interrupt Source}
The interrupt source is modeled as an object of class \verb|IntSrc|,
with attributes \verb|cycle| denoting the interrupt cycle $T$, and
\verb|val| for the value which will decrease from $T$ to $0$ while
time advances:
\begin{verbatim}
  class IntSrc | val : Time, cycle : Time .
\end{verbatim}

\subsubsection{System}
The entire system in our model is a composition of the parts
introduced above, of a sort \verb|System|\footnote{Following the Maude
  convention, variables would be written in capital letters. Variable
  declarations are not shown for simplicity.}:
\begin{verbatim}
  op _____ : TaskList Timer SysTasks 
             Hardware Object ~> System [ctor] .
  mb (L T STS HW < O : IntSrc |>) : System .
\end{verbatim}
where ``\verb|~>|'' means that the function defined is a partial
function and the keyword \verb|mb| declares a membership axiom,
stating here that a term composed of a \verb|TaskList|, a
\verb|Timer|, a \verb|SysTasks|, a \verb|Hardware| and an object
instance of class \verb|IntSrc| is of sort \verb|System|.

\subsection{Interrupt Requests}
\label{ss:ir}
Interrupt requests are performed by the source exactly every cycle
$T$, when the attribute \verb|val| decreases to \verb|zero|. The
requesting is an instantaneous action, thus is modeled by the
following instantaneous conditional rule applied on \verb|System|:
\begin{verbatim}
  crl [interrupt-request] :
    (L T STS HW ISRC) 
    => (L T STS (HW).intReq reset(ISRC))
    if (ISRC).timeout .
\end{verbatim}
where the function \verb|_.timeout| examines whether the attribute
\verb|val| equals \verb|zero|, and \verb|_.intReq| sets the \verb|ir|
bit indicating there exists an interrupt request to be handled.
Then the request will wait to be handled, which is explained in
Section~\ref{ss:inthandling}.

\subsection{Task Initiation}
\label{ss:init}
Periodic tasks are initiated sequentially by $\mathit{updateStatus()}$
in Fig.~\ref{f:schedule}, which is treated as an instantaneous action
in our model. It is modeled by function \verb|updateStatus_with_|:
\begin{verbatim}
  op updateStatus_with_ : TaskList Timer 
                            -> TaskList . 
\end{verbatim}
which applies function \verb|update_with_| on individual task
sequentially to update the status (Lines~21--30 in
Fig.~\ref{f:schedule}):
\begin{verbatim}
  op update_with_ : Object Timer ~> Object .
  ceq update < O : PTask | period : T, 
                           status : ST > 
        with TIMER
      = if ST == DORMANT 
        then < O : PTask | status : READY >
        else error fi
      if TIMER rem T == 0 .
  eq update < O : PTask | status : ST > 
       with TIMER
     = if ST == RUNNING 
       then < O : PTask | status : INTERRUPT >
       else < O : PTask |> fi [otherwise] .
\end{verbatim}
with \verb|TIMER| the current value of the global variable $\mathit{timer}$.
Given a task, if \verb|TIMER|($\mathit{timer}$) can be divided by its period
\verb|T|, this task should be initiated.  In the case where the task
should be initiated, it is set \verb|READY| if its \verb|status| is
\verb|DORMANT|; otherwise that means the previous job of this task is
not complete, hence it misses its deadline, producing an
\verb|error|. In the other case where the task should not be
initiated, its \verb|status| changes only if it is \verb|RUNNING|.

We can see that \verb|updateStatus_with_| behaves the same as
$\mathit{updateStatus()}$ in Fig.~\ref{f:schedule}.

\subsection{Interrupt Handling and Task Scheduling}
\label{ss:inthandling}
When an interrupt request occurs, it may not be detected immediately
by the system. It requires the bit \verb|mask| to be cleared. Once the
request is detected, it is handled in two steps: the interrupt
handling mechanism of the hardware (such as clearing \verb|ir|,
pushing context into stack and so on), and to invoke the function
$\mathit{schedule()}$. This behavior is modeled by the following
instantaneous rewrite rule:
\begin{verbatim}
  crl [interrupt-handle] :
    SYSTEM 
    => ((SYSTEM).interrupt).startScheduling
    if (SYSTEM).existInt .
\end{verbatim}
where \verb|_.existInt| checks whether \verb|mask| is cleared
\emph{and} \verb|ir| is set. The function \verb|_.interrupt| models
the interrupt handling mechanism performed by the hardware and does
four things: (i)~clearing the bit \verb|ir|, which means the request
has been handled; (ii)~pushing the current \verb|pc| into the stack,
storing the interrupted context; (iii)~assigning \verb|scheduling| of
sort \verb|Oid| to \verb|pc|, which indicates that the system is
scheduling; and (iv)~setting the bit \verb|mask|, to mask coming
interrupt requests.

Unlike periodic tasks, even though the scheduling stage is modeled by
a \verb|Counter|, its functionality is too important to be abandoned.
We divide the behaviors of scheduling into three parts.  The first
part contains its timed behaviors. This part is modeled by regarding
scheduling as a system task of sort \verb|SysTask|. Modeling timed
behaviors of tasks is explained in Section~\ref{ss:timedbehavior}.
The other two parts together define its functionality. The second part
corresponds to Lines~3--4 in Fig.~\ref{f:schedule}. It updates the
status of $\mathit{taskList}$ and increases $\mathit{timer}$ by
$1$. This part is modeled by function \verb|_.startScheduling| which
applies instantaneously at the beginning of scheduling, as shown in
rule \verb|interrupt-handle|:
\begin{verbatim}
  op _.startScheduling : System -> System .
  eq (L T STS HW ISRC).startScheduling 
     = ((updateStatus L with T) 
        inc(T) STS HW ISRC) .
\end{verbatim}
The third part corresponds to Lines~6--11, searching the first
should-be-run periodic task and setting it to execute. It is modeled
by function \verb|_.finishScheduling|, which applies instantaneously
at the end of scheduling:
\begin{verbatim}
  op _.finishScheduling : System -> System .
  eq (L T STS HW ISRC).finishScheduling
     = (L T (finish scheduling in STS) 
        HW ISRC).run1stTask .
\end{verbatim}
where \verb|finish_in_| resets the counter of task \verb|scheduling|,
and \verb|_.run1stTask| models Lines~6--11,
searching the task with highest priority that has status
\verb|INTERRUPT| or \verb|READY| then performing an \emph{interrupt
  return} or executing it, respectively.

When the execution time of the system task \verb|scheduling| reaches
its computation requirement, scheduling is finished. We model this
instantaneous action with the following rule:
\begin{verbatim}
  crl [scheduling-finish] :
    SYSTEM => (SYSTEM).finishScheduling
    if SYSTEM := (L T STS HW ISRC) 
       /\ (SYSTEM).running == scheduling 
       /\ scheduling isComplete?in STS .
\end{verbatim}
where function \verb|_.running| returns the current \verb|pc| value of
the system, and \verb|_isComplete?in_| checks whether the execution
time of the task reaches its computation requirement.

Similar to scheduling, the switching stage is also divided into timed
behaviors of \verb|switching| and its functionality. \verb|switching|
starts when the running periodic task is complete, and finishes when
itself is so. Two similar instantaneous rules \verb|switching-start|
and \verb|switching-finish| are defined to model the functionality of
switching.

\subsection{Timed Behaviors of the System}
\label{ss:timedbehavior}
Timed behaviors of the system consist of two parts: the execution of
tasks and the execution of the interrupt source. Both are modeled
together by the  single \emph{standard} tick
rule\footnote{The keyword \texttt{nonexec} should be given to allow
  the Real-Time Maude engine to apply the rule with some
  strategies.}~\cite{DBLP:journals/entcs/OlveczkyM07a}:
\begin{verbatim}
  crl [tick]:
    {SYSTEM} => {delta(SYSTEM, R)} in time R 
    if R le mte(SYSTEM) [nonexec] .
\end{verbatim}
where \verb|delta| defines effects of time elapse on the system, and
\verb|mte| denotes the \emph{m}aximum amount of \emph{t}ime allowed to
\emph{e}lapse from the current state until an instantaneous action
\emph{must} happen. In fact, the key to modeling timed behaviors is to
define \verb|delta| and \verb|mte|. Note that the variable \verb|R| is
\emph{continuous} with respect to the specific time
domain\footnote{Real-Time Maude contains built-in modules to define
  the time domain to be natural numbers and rational numbers,
  specifying \emph{discrete} time domains and \emph{dense} time
  domains, respectively.}  that we choose to instantiate our model on,
which is different from timed automata that discretize dense time by
defining ``clock region''.

Time affects the system by advancing both the running task, whose
$\mathit{ID}$ is loaded at \verb|pc|, and the interrupt source
simultaneously.  While time elapses, \verb|cnt| of the former
increases and \verb|val| of the latter decreases, respectively:
\begin{verbatim}
  ceq delta((L T STS HW ISRC), R)
      = (deltaTask(ID, L, R) 
         T STS HW (deltaIS(ISRC, R)))
      if ID := (HW).getPc /\ ID :: MaybeNat .
\end{verbatim}
where the last condition indicates that \verb|ID| is of sort
\verb|MaybeNat|. Due to similarity, we omit details for the case where
\verb|ID| is of sort \verb|Oid| and \verb|deltaTask| applies on
\verb|STS| instead of \verb|L|.

\verb|mte| depends on when the next instantaneous transition must
perform. Therefore, it is decided by three arguments: the remaining
time to complete the running task, the remaining time to request the
next interrupt, and whether or not there exists an interrupt request
detected for the moment:
\begin{verbatim}
  ceq mte(L T STS HW ISRC)
      = minimum(mteTask(ID, L),
                mteIS(ISRC), mteIr(HW))
      if ID := (HW).getPc /\ ID :: MaybeNat .
\end{verbatim}
where \verb|mteIr| returns \verb|zero| if there exists an interrupt
request detected in the system, or \verb|INF| which represents
\emph{infinity} otherwise. The case where \verb|ID| is of sort
\verb|Oid| is similar.

\section{Formal Verification}
\label{s:verification}
In this section, we analyze our model of the RMS implementation within
different realistic scenarios.  Notice that from any (reasonable)
given initial state, the number of reachable states is finite, but may
be unknown, thanks to the upper bound given to $\mathit{timer}$, which
provides the potential for applying the untimed model checker.

\subsection{Properties}
We take two properties into account in this paper: schedulability and
correctness. By schedulability, we examine whether a given task set is
schedulable by the implementation. By correctness, we verify whether
the implementation schedules periodic tasks exactly with respect to
the RMS algorithm.

To verify the schedulability of a given set of periodic tasks, we
define an atomic proposition \verb|taskTimeout| to hold if there
exists an \verb|error| in $\mathit{taskList}$ of the current state,
that is, some task misses its deadline:
\begin{verbatim}
  op taskTimeout : -> Prop [ctor] .
  eq {L T STS HW ISRC} |= taskTimeout 
     = containError(L) .
\end{verbatim}
where \verb|containError| returns \verb|true| if there is an
\verb|error| existing in \verb|L|. Then schedulability can be
formalized as the temporal logic formula: \verb|[](~taskTimeout)|,
expressing that the proposition \verb|taskTimeout| is always false. As
the property is not \emph{clock-related}, given an initial state
\verb|init|, the following untimed model checking command returns
\verb|true| if the schedulability property holds with no time limit;
otherwise a trace showing a counterexample is provided:
\begin{verbatim}
  (mc init |=u [](~taskTimeout) .)
\end{verbatim}

Another important objective is to verify the correctness of the
implementation.  The atomic proposition \verb|correct| is hence
defined to hold if the running periodic task is the one requested to
be executed with the highest priority:
\begin{verbatim}
  op correct : -> Prop [ctor] .
  ceq {L T STS HW ISRC} |= correct
      = if ID :: MaybeNat then shouldRun(ID, L)
        else true fi
      if ID := (HW).getPc .
\end{verbatim}
where \verb|shouldRun(ID, L)| returns \verb|true| if the task
identified by \verb|ID|, probably \verb|none|, is the one possessing
the highest priority among those whose status is not
\verb|DORMANT|. Note that during verification, we do not care the
behaviors after some task misses its deadline. Therefore, the
correctness property is formalized by the temporal logic formula:
\verb|([]correct)\/(correct U taskTimeout)|, stating that
\verb|correct| is always true, or is true until \verb|taskTimeout|
holds. It can be verified by the following untimed model checking
command provided an initial state \verb|init|:
\begin{verbatim}
  (mc init |=u ([]correct) 
               \/ (correct U taskTimeout) .)
\end{verbatim}

\subsection{Scenarios}
\label{ss:results}
We use the following setting for our verification, which is from the 
statistics provided by our industrial partner:
\begin{itemize}
\item The interrupt cycle $T$ is $5ms$.
\item The scheduling time is $38{\mu}s$, switching time is $20{\mu}s$.
\item The initial state is with empty stack, empty \verb|pc|, cleared 
\verb|mask| and cleared \verb|ir|.
\end{itemize}

We have analyzed our model in ten different scenarios, including both
realistic ones provided by our industrial partner and experimental
ones designed by ourselves, four of them are described below:
\begin{itemize}
\item Scenario~(i) with 2 tasks $\tau_1$ and $\tau_2$: $T_1=5ms$,
  $C_1=3ms$, $T_2=25ms$, $C_2=7ms$.
\item Scenario~(ii) with 2 tasks $\tau_1$ and $\tau_2$: $T_1=5ms$,
  $C_1=2ms$, $T_2=25ms$, $C_2=2.3ms$.
\item Scenario~(iii) with 3 tasks $\tau_1$, $\tau_2$, $\tau_3$:
  $T_1=5ms$, $C_1=2.7ms$, $T_2=10ms$, $C_2=2ms$, $T_3=25ms$,
  $C_3=3ms$.
\item Scenario~(iv) with 3 tasks $\tau_1$, $\tau_2$, $\tau_3$:
  $T_1=5ms$, $C_1=2.5ms$, $T_2=10ms$, $C_2=1.5ms$,
  $T_3=15ms$, $C_3=4.5ms$.
\end{itemize}
Note that thanks to the expressiveness of Real-Time Maude, we
only need to define an initial state of sort \verb|System| to specify
a given task set. No necessity to modify the model is needed.

\begin{figure*}[!t]
\centering
\includegraphics{FIG3_15-TIE-3480.pdf}
\caption{A counterexample of schedulability in Scenario~(iv). The first
  job instance of $\tau_3$ misses its deadline at time $15ms$, which
  is the initiation time for the second job instance of $\tau_3$.}
\label{f:counterexample}
\end{figure*}

Instantiating our model on dense time domain and choosing the
\emph{maximal time sampling strategy}, the results of the model
checking show that correctness property holds in all scenarios.
Schedulability property holds in Scenarios~(i--iii), but fails in
Scenario~(iv).  One counterexample of the schedulability within
Scenario~(iv), returned by the model checking command, is pictured in
Fig.~\ref{f:counterexample}, where $\tau_3$ misses its deadline at
time $15ms$.

The results above have demonstrated that our approach is capable of
handling realistic industrial systems. However, to further exam the
efficiency of our approach, we also apply our method to larger test
scenarios. Test scenarios are generated randomly. We verify the
schedulability of them under the above setting on an Intel Core 2 Quad
Q9550, 2.83GHz, 4-core machine with 8GB RAM running 64-bit Ubuntu
15.04. Among the 50 generated test scenarios with 5 tasks, we find
that the execution time of the model checking command for each
scenario varies from about $300ms$ up to a timeout, which is set to
$90$ minutes. This is because the efficiency of model checking depends
on the scale of the state space. And the scale is further positively
correlated with $mn$, where $m$ is the upper bound of $\mathit{timer}$
and $n$ is the number of periodic tasks. By the generated test
scenarios, it turns out that the model checking command for
schedulability in our model is able to handle scenarios, where $mn$ is
up to about $10^6$, in an acceptable period of time say $90$ minutes.

\subsection{Evaluation}
We now show in this section that our results are both sound and
complete.

An analysis method is called \emph{sound} if any counterexample found
using such a method is a real counterexample of the question, and
\emph{complete} if the fact that no counterexample can be found using
such a method implies no counterexample exists for the question in
analysis. The soundness of our results is trivial to check, simply by
examining the counterexamples found. For instance, the counterexample
shown in Fig.~\ref{f:counterexample} is a real counterexample,
implying that the result for schedulability of Scenario~(iv) is
sound. But this is not the case for completeness, since we choose
instantiating our model on dense time domain to make it more real but
giving rise to an infinite state space which is unfeasible to exhaust.

In general, completeness of untimed model checking cannot be achieved
for any systems, any time sampling strategies and any
properties. However, \"Olveczky and Meseguer proved the completeness
of untimed temporal logic model checking, under the maximal time
sampling strategy, for a large class of real-time systems possessing a
set of ``good'' properties that is called \emph{time-robustness}, and
for a set of ``good'' LTL formulae constructed by
\emph{tick-invariant} propositions\footnote{We avoid introducing the
  definitions of time-robustness and tick-invariance, due to the
  requirements of extra rewriting logic
  background.}~\cite{DBLP:journals/entcs/OlveczkyM07a}:
\begin{theorem}[\cite{DBLP:journals/entcs/OlveczkyM07a}]
\label{t:completeness}
Given a time-robust real-time rewrite theory $\mathcal{R}$, a set $AP$
of tick-invariant atomic propositions, an LTL formula $\Phi$
(excluding the \emph{next} operator $\bigcirc$) whose atomic
propositions are contained in $AP$. The untimed temporal logic model
checking verifying $\Phi$ is \emph{complete} under the maximal time
sampling strategy.
\end{theorem}

Therefore, we achieve the following theorem, showing that the results
in Section~\ref{ss:results} are complete.
\begin{theorem}
\label{t:main}
Our approach using untimed model checking to verify schedulability
and correctness of our model is complete.
\end{theorem}
\begin{IEEEproof}
By showing that our model is time-robust and that the two defined
atomic propositions -- \verb|taskTimeout| and \verb|correct| -- are
tick-invariant, then applying Theorem~\ref{t:completeness}. For a more
detailed proof, see the Appendix.
\end{IEEEproof}

\section{Related Work}
\label{s:relate}
In this section we discuss our results with related
work in three directions.

Considering schedulability test, Liu and
Layland~\cite{DBLP:journals/jacm/LiuL73} gave the famous sufficient
condition that a set of periodic tasks is schedulable with respect to
RMS if $\Sigma^n_{i=1} C_i/T_i \le n(2^{1/n}-1)$ holds. Then a more
sufficient condition, known as \emph{Hyperbolic Bound}, which has the
same complexity as Liu and Layland's, was proposed
in~\cite{DBLP:journals/tc/BiniBB03}. On the other hand, necessary and
sufficient conditions for schedulability were derived independently
in~\cite{DBLP:journals/rts/SpruntSL89} and~\cite{audsley1993deadline},
requiring more sophisticated analysis on the task set. Nevertheless,
all these results take no overhead into account, being not as
realistic as ours. Katcher et al.\ did consider overhead in their
schedulability analysis, under several models based on different kinds
of popular implementations~\cite{DBLP:journals/tse/KatcherAS93}.
However, our target implementation is not in their scope. Furthermore,
compared with those theoretical analyses, our approach based on formal
modeling and verification has three advantages. One is that if our
schedulability test answers ``no'', it returns at the same time a real
counterexample, which is able to guide our engineer to adjust the
design, by changing either the priorities or even the scheduling
algorithm. The second is that, when a fresh scheduling strategy is
applied, our analysis can be adjusted only by modifying the model,
while theoretical approaches may need thorough analyses and reasoning.
The last one is that, considering overhead and details of hardware
does introduce some kind of non-determinism into the model. For
example, in our model, if the running task is complete \emph{right} at
the time when an interrupt request occurs, two different behaviors are
possible: (i)~the system performs task switching, during which the
interrupt request is masked, hence the scheduling and initiation of
tasks may be delayed; (ii)~the system answers the interrupt request
immediately, the switching being delayed. Tackling such
non-determinism via theoretical analyses seems more complicated than
our automatic approach.

\cite{TianD2011} and~\cite{DBLP:conf/iceccs/CuiDT14} also made use of
model checking to analyze the RMS algorithm along the same line but
with different languages and tools. The RMS algorithm is investigated
using an extension of SPIN~\cite{DBLP:journals/tse/Holzmann97}
in~\cite{TianD2011}, while the logic programming language
TMSVL~\cite{DBLP:conf/icfem/HanDW12} and its model checker are applied
in~\cite{DBLP:conf/iceccs/CuiDT14}. The work presented
in~\cite{TianD2011} and~\cite{DBLP:conf/iceccs/CuiDT14} has two main
differences with ours. The first one, which is also the motivational
one, is that they considered only the ideal setting of the algorithm
as in~\cite{DBLP:journals/jacm/LiuL73}, instead of an implementation
which contains much more complex details. In fact, our model of the
implementation can easily degenerate to a model of the RMS algorithm,
if we let the times for scheduling and switching be zero. The other
difference is that when adding a new task into the task set, the
models in~\cite{TianD2011} and~\cite{DBLP:conf/iceccs/CuiDT14} should
be modified by explicitly defining a sub-model of the new task and its
behaviors. In particular, the scheduling part of the model
in~\cite{TianD2011} needs adjustments as well to include a new
task. Nevertheless, a new task set is specified in our approach merely
by giving a new initial state, without necessity to modify the model,
as already emphasized in Section~\ref{ss:results}. On the other hand,
\cite{TianD2011} used a discrete time domain in the model, while we
employ a generic time domain, which is flexible to be instantiated on
either discrete or dense ones. In~\cite{DBLP:conf/iceccs/CuiDT14}, the
time domain is dense, and a modeling strategy like our maximal time
sampling strategy is applied to reduce the state space. However, a
completeness statement like Theorem~\ref{t:main} was not given
in~\cite{DBLP:conf/iceccs/CuiDT14}.

Finally, Maude and Real-Time Maude have been successfully applied on
large numbers of applications~\cite{DBLP:journals/jlp/Meseguer12},
especially on communication protocols, real-time and cyber-physical
systems. But few results are achieved on scheduling problems. RMS is
investigated using Real-Time Maude for the first time.

\section{Conclusion}
\label{s:conclusion}
We have formally modeled a realistic implementation of RMS using
Real-Time Maude, a modeling language based on rewriting logic. By
taking into account the overhead of scheduling and switching, and by
modeling some mechanism of the hardware, our model contains sufficient
details to be analyzed for behaviors of the real target
system. Two important properties -- schedulability and correctness --
are verified by model checking on our model, within several key
scenarios. We demonstrate the soundness and completeness of our
results.

\appendix[Proof of Theorem~\ref{t:main}]
\label{a:proof}
\setcounter{theorem}{0}
\renewcommand{\thetheorem}{\thesection.\arabic{theorem}}
\newcommand{\lrps}[2]{\rightarrow^{#1}_{#2}}

We show here a detailed proof of Theorem~\ref{t:main}.

Some more preliminaries from~\cite{DBLP:journals/entcs/OlveczkyM07a}
are needed. A term $t$ is called \emph{ground} if it contains no
variables. For a set $P\subseteq AP$ of atomic propositions and ground
terms $t,t'$, we write $t\simeq_P t'$ (or $t\simeq t'$ for simplicity
when $P$ is implicit) if $t$ and $t'$ satisfy exactly the same set of
propositions from $P$.

Time-robustness is a set of properties expected from a well-behaved
real-time rewrite theory. We avoid introducing the accurate definition
of time-robustness, but instead give the following lemma to prove
time-robustness:
\begin{lemma}[\cite{DBLP:journals/entcs/OlveczkyM07a}]
\label{l:timerobustness}
Let $\mathcal{R}$ be an object-oriented specification with a standard
tick rule, and let the infinity element \verb|INF| be the only element
in the time domain which is not a normal time value.  Then
$\mathcal{R}$ is time-robust if the following conditions are satisfied
for all appropriate ground terms $t$ and $r,r'$:
\begin{enumerate}
\item
[(i)] \verb|mte(delta(|$t$\verb|,|$r$\verb|))| $=$
\verb|mte(|$t$\verb|) monus |$r$, for all $r\le$
\verb|mte(|$t$\verb|)|, where \verb|monus| is the built-in minus
operation defined on sort \verb|Time|;

\item
[(ii)] \verb|delta(|$t$\verb|,|$0$\verb|)| $= t$;

\item
[(iii)] \verb|delta(delta(|$t$\verb|,|$r$\verb|),|$r'$\verb|)| $=$
\verb|delta(|$t$\verb|,|$r+r'$\verb|)|, for $r+r'\le$
\verb|mte(|$t$\verb|)|;

\item
[(iv)] \verb|mte(|$\sigma(l)$\verb|)|$= 0$ for each ground instance
$\sigma(l)$ of each left-hand side $l$ of an instantaneous rewrite
rule.
\end{enumerate}
\end{lemma}

Each one-step rewrite can be categorized and then tick-invariance can
be defined:
\begin{definition}[\cite{DBLP:journals/entcs/OlveczkyM07a}]
A one-step rewrite $t\lrps{r}{1}t'$ using a tick rule and having
duration $r$ is:
\begin{itemize}
\item[--] a \emph{maximal tick step}, written $t\lrps{r}{max}t'$, if
  there is no time value $r'>r$ such that $t\lrps{r'}{1}t''$ for some
  $t''$;
\item[--] an \emph{$\infty$ tick step}, written $t\lrps{r}{\infty}t'$,
  if for each time value $r'>0$, there is a tick rewrite step
  $t\lrps{r'}{1}t''$; and
\item[--] a \emph{non-maximal tick step} if there is a maximal tick
  step $t\lrps{r'}{max}t''$ for $r'>r$.
\end{itemize}
\end{definition}
\begin{definition}[\cite{DBLP:journals/entcs/OlveczkyM07a}]
  A time-robust specification $\mathcal{R}$ is \emph{tick-invariant} with
  respect to a set $P$ of propositions if and only if $t\simeq_P t'$
  holds for each non-maximal or $\infty$ tick step $t\lrps{r}{}t'$.
\end{definition}

The following lemma is needed to prove the tick-invariance of our
defined propositions:
\begin{lemma}[\cite{DBLP:journals/entcs/OlveczkyM07a}]
\label{l:tickinv}
Let $\mathcal{R}$ be a time-robust object-oriented specification with a
standard tick rule, and let the infinity element \verb|INF| be the
only element in the time domain which is not a normal time value.
Then $\mathcal{R}$ is tick-invariant with respect to a set $P$ of atomic
propositions if \verb|{|$t$\verb|}| $\simeq_P$
\verb|{delta(|$t$\verb|,|$r$\verb|)}| for all $t,r$ with $r<$
\verb|mte(|$t$\verb|)|.
\end{lemma}

\newcommand{\mteTask}[2]{\texttt{mteTask(}#1\texttt{,}#2\texttt{)}}
\newcommand{\deltaTask}[3]{\texttt{deltaTask(}#1\texttt{,}#2\texttt{,}#3\texttt{)}}
\newcommand{\mteIS}[1]{\texttt{mteIS(}#1\texttt{)}}
\newcommand{\deltaIS}[2]{\texttt{deltaIS(}#1\texttt{,}#2\texttt{)}}
\newcommand{\IntSrc}[3]{\texttt{<}#1\texttt{:IntSrc|val:}#2\texttt{,cycle:}#3\texttt{>}}
\newcommand{\mteIr}[1]{\texttt{mteIr(}#1\texttt{)}}
\newcommand{\mteS}[1]{\texttt{mte(}#1\texttt{)}}
\newcommand{\deltaS}[2]{\texttt{delta(}#1\texttt{,}#2\texttt{)}}

Several auxiliary lemmas are proved:
\begin{lemma}
\label{l:auxtask}
Given $\mathit{ID}$ of sort \verb|MaybeNat| and $L$ of sort
\verb|TaskList|, $\mteTask{\mathit{ID}}{\deltaTask{\mathit{ID}}{L}{r}}
= \mteTask{\mathit{ID}}{L}\; \verb|monus|\;r$ for all
$r\le\mteTask{\mathit{ID}}{L}$.
\end{lemma}
\begin{IEEEproof}
If $\mathit{ID}=\verb|none|$, the case is trivial. By definition,
$\mteTask{\texttt{none}}{\deltaTask{\texttt{none}}{L}{r}} =
\mteTask{\texttt{none}}{L} = \verb|INF| =
\mteTask{\texttt{none}}{L}\\ \verb|monus|\; r$. Otherwise,
$\mathit{ID}= \verb|some|~N$ with $N$ of sort \verb|Nat|. Assume that
the \verb|cnt| of the $N$th task in $L$ is
\verb|[|$r_e$\verb|/|$C$\verb|]|. Then by definition,
$\deltaTask{\mathit{ID}}{L}{r}=L'$, where the
$N$th task in $L'$ has \verb|cnt| value being
\verb|[|$(r_e+r)$\verb|/|$C$\verb|]|. Hence,
\begin{IEEEeqnarray*}{Cl}
  & \mteTask{\mathit{ID}}{\deltaTask{\mathit{ID}}{L}{r}}
\\  
= & \mteTask{\mathit{ID}}{L'}
\\
= & C~\verb|monus|~(r_e+r) 
= (C~\verb|monus|~r_e)~\verb|monus|~r 
\\
= & \mteTask{\mathit{ID}}{L}~\verb|monus|~r\;. \hfill\IEEEQEDhere
\end{IEEEeqnarray*}
\end{IEEEproof}

\begin{lemma}
\label{l:auxis}
Given $\mathit{ISRC}$ of class \verb|IntSrc| representing an
reasonable state of interrupt source in our model,
$\mteIS{\deltaIS{\mathit{ISRC}}{r}} = \mteIS{\mathit{ISRC}}\;
\verb|monus|\\ r$ for all $r\le\mteIS{\mathit{ISRC}}$.
\end{lemma}
\begin{IEEEproof}
A reasonable $\mathit{ISRC}$ must be of the form
$\IntSrc{O}{v}{T}$ with $v\le T$. Thus,
\begin{IEEEeqnarray*}{Cl}
  & \mteIS{\deltaIS{\mathit{ISRC}}{r}}
\\  
= & \mteIS{\IntSrc{O}{(v~\texttt{monus}~r)}{T}}
\\
= & v~\texttt{monus}~r
= \mteIS{\mathit{ISRC}}~\texttt{monus}~r\;. \hfill\IEEEQEDhere
\end{IEEEeqnarray*}
\end{IEEEproof}

\begin{lemma}
\label{l:auxhw}
Given $\mathit{HW}$ of sort \verb|Hardware|, for all
$r\le\mteIr{\mathit{HW}}$, $\mteIr{\mathit{HW}} =
\mteIr{\mathit{HW}}\; \verb|monus|\; r$.
\end{lemma}
\begin{IEEEproof}
  It concludes by discussing on whether there exists an interrupt
  request detected in $\mathit{HW}$.
\end{IEEEproof}

Now we can present the detailed proof of Theorem~\ref{t:main}:
\begin{IEEEproof}[Proof of Theorem~\ref{t:main}]
We first prove the time-robustness of our model by
Lemma~\ref{l:timerobustness}.  Instantiated on the built-in dense time
domain \verb|POSRAT-TIME-DOMAIN-WITH-INF|, our model has \verb|INF| as
the only element which is not a normal time value. As presented in
Section~\ref{ss:timedbehavior}, only a single standard tick rule is
defined. Hence our model is time-robust by
Lemma~\ref{l:timerobustness} provided conditions (i--iv) hold.

Condition~(i). We must prove \verb|mte(delta(|$s$\verb|,|$r$\verb|))|
$=$ \verb|mte(|$s$\verb|)| \verb|monus| $r$ for all $r\le$
\verb|mte(|$s$\verb|)|, with $s$ being a system state of sort
\verb|System|. Let $s$ be of the form
$(L~T~\mathit{STS}~\mathit{HW}~\mathit{ISRC})$ and
$\mathit{ID}=\verb|(|\mathit{HW}\verb|).getPc|$. We only consider the
case $\mathit{ID}$\verb|::MaybeNat| in detail, while the other case
$\mathit{ID}$\verb|::Oid| is similar. By definitions of \verb|mte| and
\verb|delta|,
\begin{IEEEeqnarray*}{Cl}
  & \mteS{\deltaS{s}{r}}
\\  
= & \mteS{\deltaTask{\mathit{ID}}{L}{r}
\\
  & \;\;\;\;\;\;\;\;\;T~\mathit{STS}~\mathit{HW}~\deltaIS{\mathit{ISRC}}{r}}
\\
= & \verb|minimum(|\mteTask{\mathit{ID}}{\deltaTask{\mathit{ID}}{L}{r}}\verb|,|
\\
  & \verb|        |\mteIS{\deltaIS{\mathit{ISRC}}{r}}\verb|,|
\\
  & \verb|        |\mteIr{\mathit{HW}}\verb|)|\;.
\end{IEEEeqnarray*}
By Lemmas~\ref{l:auxtask},~\ref{l:auxis} and~\ref{l:auxhw}, it can be
further reduced:
\begin{IEEEeqnarray*}{Cl}
  & \mteS{\deltaS{s}{r}}
\\  
= & \verb|minimum(|\mteTask{\mathit{ID}}{L}~\verb|monus|~r~\verb|,|
\\
  & \verb|        |\mteIS{\mathit{ISRC}}~\verb|monus|~r~\verb|,|
\\
  & \verb|        |\mteIr{\mathit{HW}}~\verb|monus|~r~\verb|)|
\\
= & \verb|minimum(|\mteTask{\mathit{ID}}{L}\verb|,|
\\
  & \verb|        |\mteIS{\mathit{ISRC}}\verb|,|
\\
  & \verb|        |\mteIr{\mathit{HW}}\verb|)|~\verb|monus|~r
\\
= & \mteS{s}~\verb|monus|~r\;.
\end{IEEEeqnarray*}

Condition~(ii) follows from the fact that $r+0=r$ and
$r~\verb|monus|~0=r$ for all $r$.

Condition~(iii). We must prove $\deltaS{\deltaS{s}{r}}{r'}\\ =
\deltaS{s}{r+r'}$ for all $r+r'\le\mteS{s}$, with $s$ being a system
state of sort \verb|System|. Using the same notations as in~(i), we
only consider the case $\mathit{ID}$\verb|::MaybeNat| in detail. By
definition, the left side of the equation
\begin{IEEEeqnarray*}{Cl}
  & \deltaS{\deltaS{s}{r}}{r'}
\\  
= & \verb|(|\deltaTask{\mathit{ID}}{\deltaTask{\mathit{ID}}{L}{r}}{r'}
\\
  & \verb| |~T~\mathit{STS}~\mathit{HW}~\deltaIS{\deltaIS{\mathit{ISRC}}{r}}{r'}\verb|)|~\mbox{,}
\end{IEEEeqnarray*}
and the right side of the equation
\begin{IEEEeqnarray*}{Cl}
  & \deltaS{s}{r+r'}
\\  
= & \verb|(|\deltaTask{\mathit{ID}}{L}{r+r'}
\\
  & \verb| |~T~\mathit{STS}~\mathit{HW}~\deltaIS{\mathit{ISRC}}{r+r'}\verb|)|~\mbox{.}
\end{IEEEeqnarray*}
$\deltaTask{\mathit{ID}}{\deltaTask{\mathit{ID}}{L}{r}}{r'}
=\deltaTask{\mathit{ID}}{L}{r+r'}$ holds thanks to the associativity
of $+$, while
$\deltaIS{\deltaIS{\mathit{ISRC}}{r}}{r'}\!=\!\deltaIS{\mathit{ISRC}}{r+r'}$
holds since $(v\; \verb|monus|\; r)\; \verb|monus|\; r'\\ = v\;
\verb|monus|\; (r+r')$ with $v$ of sort \verb|Time|. Thus (iii) holds.

Condition~(iv). We show that \verb|mte| of each instance of the
left-hand side of any instantaneous rule is $0$. For example,
considering the rule \verb|interrupt-request|, with its
condition \verb|(|$\mathit{ISRC}$\verb|).timeout|, we know that \verb|val|
of $\mathit{ISRC}$ equals $0$. Therefore,
\begin{IEEEeqnarray*}{Cl}
  & \mteS{L~T~\mathit{STS}~\mathit{HW}~\mathit{ISRC}}
\\  
= & \verb|minimum(|\mteTask{\mathit{ID}}{L}\verb|,|
\\
  & \verb|        |\mteIS{\mathit{ISRC}}\verb|,| \mteIr{\mathit{HW}}\verb|)|
\\
= & \verb|minimum(|\mteTask{\mathit{ID}}{L}\verb|,|~0~\verb|,| \mteIr{\mathit{HW}}\verb|)|
\\
= & 0~\mbox{.}
\end{IEEEeqnarray*}
The other rules can be similarly proved with their conditions. Hence
our model is time-robust by Lemma~\ref{l:timerobustness}.

Finally we prove the tick-invariance of the propositions used to
analyze our model, i.e.\ \verb|taskTimeout| and \verb|correct|. By
Lemma~\ref{l:tickinv}, we must prove
$\verb|{|s\verb|}|\simeq_P\verb|{|\deltaS{s}{r}\verb|}|$ with $s$ of
sort \verb|System| and $r<\mteS{s}$, that is, applying a tick rule
advancing $r$ time units will not change the value of each
proposition. Let $s$ be $(L~T~\mathit{STS}~\mathit{HW}~\mathit{ISRC})$
and $\verb|(|\mathit{HW}\verb|).getPc|=\mathit{ID}$.
\begin{itemize}
\item \verb|taskTimeout| holds if and only if $L$ contains
  \verb|error|s. Since \verb|delta| does not produce or eliminate
  \verb|error| in $L$, \verb|taskTimeout| holds in $s$ if and only if
  \verb|taskTimeout| holds in $\deltaS{s}{r}$ for any $r<\mteS{s}$,
  implying our model is tick-invariant with respect to
  \verb|taskTimeout|.
\item The value of \verb|correct| depends on $\mathit{ID}$ and
  status of each task in $L$. Similarly, \verb|delta| does not change
  $\mathit{ID}$ or status of any task in $L$, hence \verb|correct|
  holds in $s$ if and only if \verb|correct| holds in $\deltaS{s}{r}$
  for any $r<\mteS{s}$. It concludes the tick-invariance of
  \verb|correct|.
\end{itemize}

Therefore, by Theorem~\ref{t:completeness}, our approach using untimed
model checking to verify schedulability and correctness is complete.
\end{IEEEproof}

%% References

% Generated by IEEEtran.bst, version: 1.12 (2007/01/11)
\begin{thebibliography}{10}
\providecommand{\url}[1]{#1}
\csname url@samestyle\endcsname
\providecommand{\newblock}{\relax}
\providecommand{\bibinfo}[2]{#2}
\providecommand{\BIBentrySTDinterwordspacing}{\spaceskip=0pt\relax}
\providecommand{\BIBentryALTinterwordstretchfactor}{4}
\providecommand{\BIBentryALTinterwordspacing}{\spaceskip=\fontdimen2\font plus
\BIBentryALTinterwordstretchfactor\fontdimen3\font minus
  \fontdimen4\font\relax}
\providecommand{\BIBforeignlanguage}[2]{{%
\expandafter\ifx\csname l@#1\endcsname\relax
\typeout{** WARNING: IEEEtran.bst: No hyphenation pattern has been}%
\typeout{** loaded for the language `#1'. Using the pattern for}%
\typeout{** the default language instead.}%
\else
\language=\csname l@#1\endcsname
\fi
#2}}
\providecommand{\BIBdecl}{\relax}
\BIBdecl

\bibitem{DBLP:journals/jacm/LiuL73}
C.~L. Liu and J.~W. Layland, ``Scheduling algorithms for multiprogramming in a
  hard-real-time environment,'' \emph{J. {ACM}}, vol.~20,
  \href{http://dx.doi.org/10.1145/321738.321743}{DOI 10.1145/321738.321743},
  no.~1, pp. 46--61, 1973.

\bibitem{DBLP:conf/rtss/LehoczkySS87}
J.~P. Lehoczky, L.~Sha, and J.~K. Strosnider, ``Enhanced aperiodic
  responsiveness in hard real-time environments,'' in \emph{Proceedings of the
  8th {IEEE} Real-Time Systems Symposium {(RTSS} '87)}, 1987, pp. 261--270.

\bibitem{DBLP:journals/rts/SpruntSL89}
B.~Sprunt, L.~Sha, and J.~P. Lehoczky, ``Aperiodic task scheduling for hard
  real-time systems,'' \emph{Real-Time Systems}, vol.~1,
  \href{http://dx.doi.org/10.1007/BF02341920}{DOI 10.1007/BF02341920}, no.~1,
  pp. 27--60, 1989.

\bibitem{DBLP:conf/rtss/LehoczkyR92}
J.~P. Lehoczky and S.~Ramos{-}Thuel, ``An optimal algorithm for scheduling
  soft-aperiodic tasks in fixed-priority preemptive systems,'' in
  \emph{Proceedings of the Real-Time Systems Symposium},
  \href{http://dx.doi.org/10.1109/REAL.1992.242671}{DOI
  10.1109/REAL.1992.242671}, 1992, pp. 110--123.

\bibitem{DBLP:journals/tc/StrosniderLS95}
J.~K. Strosnider, J.~P. Lehoczky, and L.~Sha, ``The deferrable server algorithm
  for enhanced aperiodic responsiveness in hard real-time environments,''
  \emph{{IEEE} Trans. Comput.}, vol.~44,
  \href{http://dx.doi.org/10.1109/12.368008}{DOI 10.1109/12.368008}, no.~1, pp.
  73--91, 1995.

\bibitem{DBLP:journals/pe/LeungW82}
J.~Y. Leung and J.~Whitehead, ``On the complexity of fixed-priority scheduling
  of periodic, real-time tasks,'' \emph{Perform. Eval.}, vol.~2,
  \href{http://dx.doi.org/10.1016/0166-5316(82)90024-4}{DOI
  10.1016/0166-5316(82)90024-4}, no.~4, pp. 237--250, 1982.

\bibitem{audsley1993deadline}
N.~Audsley, A.~Burns, and A.~Wellings, ``Deadline monotonic scheduling theory
  and application,'' \emph{Control Engineering Practice}, vol.~1, no.~1, pp.
  71--78, 1993.

\bibitem{DBLP:journals/tc/ShaRL90}
L.~Sha, R.~Rajkumar, and J.~P. Lehoczky, ``Priority inheritance protocols: An
  approach to real-time synchronization,'' \emph{{IEEE} Trans. Comput.},
  vol.~39, \href{http://dx.doi.org/10.1109/12.57058}{DOI 10.1109/12.57058},
  no.~9, pp. 1175--1185, 1990.

\bibitem{dhall1978real}
S.~K. Dhall and C.~Liu, ``On a real-time scheduling problem,'' \emph{Operations
  Research}, vol.~26, no.~1, pp. 127--140, 1978.

\bibitem{DBLP:journals/rts/LopezGDG03}
J.~M. L{\'{o}}pez, M.~Garc{\'{\i}}a, J.~L. D{\'{\i}}az, and D.~F.
  Garc{\'{\i}}a, ``Utilization bounds for multiprocessor rate-monotonic
  scheduling,'' \emph{Real-Time Systems}, vol.~24,
  \href{http://dx.doi.org/10.1023/A:1021749005009}{DOI
  10.1023/A:1021749005009}, no.~1, pp. 5--28, 2003.

\bibitem{DBLP:journals/tpds/LopezDG04}
J.~M. L{\'{o}}pez, J.~L. D{\'{\i}}az, and D.~F. Garc{\'{\i}}a, ``Minimum and
  maximum utilization bounds for multiprocessor rate monotonic scheduling,''
  \emph{{IEEE} Trans. Parallel Distrib. Syst.}, vol.~15,
  \href{http://dx.doi.org/10.1109/TPDS.2004.25}{DOI 10.1109/TPDS.2004.25},
  no.~7, pp. 642--653, 2004.

\bibitem{DBLP:journals/tc/BaruahG03}
S.~K. Baruah and J.~Goossens, ``Rate-monotonic scheduling on uniform
  multiprocessors,'' \emph{{IEEE} Trans. Comput.}, vol.~52,
  \href{http://dx.doi.org/10.1109/TC.2003.1214344}{DOI
  10.1109/TC.2003.1214344}, no.~7, pp. 966--970, 2003.

\bibitem{DBLP:journals/rts/OhS94}
Y.~Oh and S.~H. Son, ``Enhancing fault-tolerance in rate-monotonic
  scheduling,'' \emph{Real-Time Systems}, vol.~7,
  \href{http://dx.doi.org/10.1007/BF01088524}{DOI 10.1007/BF01088524}, no.~3,
  pp. 315--329, 1994.

\bibitem{DBLP:journals/rts/GhoshMMS98}
S.~Ghosh, R.~G. Melhem, D.~Moss{\'{e}}, and J.~S. Sarma, ``Fault-tolerant
  rate-monotonic scheduling,'' \emph{Real-Time Systems}, vol.~15,
  \href{http://dx.doi.org/10.1023/A:1008046012844}{DOI
  10.1023/A:1008046012844}, no.~2, pp. 149--181, 1998.

\bibitem{DBLP:journals/tpds/BertossiMR99}
A.~A. Bertossi, L.~V. Mancini, and F.~Rossini, ``Fault-tolerant rate-monotonic
  first-fit scheduling in hard-real-time systems,'' \emph{{IEEE} Trans.
  Parallel Distrib. Syst.}, vol.~10,
  \href{http://dx.doi.org/10.1109/71.798317}{DOI 10.1109/71.798317}, no.~9, pp.
  934--945, 1999.

\bibitem{DBLP:conf/rtss/LehoczkySD89}
J.~P. Lehoczky, L.~Sha, and Y.~Ding, ``The rate monotonic scheduling algorithm:
  Exact characterization and average case behavior,'' in \emph{Proceedings of
  the Real-Time Systems Symposium},
  \href{http://dx.doi.org/10.1109/REAL.1989.63567}{DOI
  10.1109/REAL.1989.63567}, 1989, pp. 166--171.

\bibitem{DBLP:conf/rtss/KuoM91}
T.~Kuo and A.~K. Mok, ``Load adjustment in adaptive real-time systems,'' in
  \emph{Proceedings of the Real-Time Systems Symposium},
  \href{http://dx.doi.org/10.1109/REAL.1991.160369}{DOI
  10.1109/REAL.1991.160369}, 1991, pp. 160--170.

\bibitem{DBLP:journals/tc/BiniBB03}
E.~Bini, G.~C. Buttazzo, and G.~M. Buttazzo, ``Rate monotonic analysis: The
  hyperbolic bound,'' \emph{{IEEE} Trans. Comput.}, vol.~52,
  \href{http://dx.doi.org/10.1109/TC.2003.1214341}{DOI
  10.1109/TC.2003.1214341}, no.~7, pp. 933--942, 2003.

\bibitem{gardner1999}
M.~K. Gardner, ``Probabilistic analysis and scheduling of critical soft
  real-time systems,'' Ph.D. dissertation, University of Illinois at
  Urbana-Champaign, 1999.

\bibitem{DBLP:journals/tie/JiangZLDSGS15}
Y.~Jiang, H.~Zhang, Z.~Li, Y.~Deng, X.~Song, M.~Gu, and J.~Sun, ``Design and
  optimization of multiclocked embedded systems using formal techniques,''
  \emph{{IEEE} Trans. Ind. Electron.}, vol.~62,
  \href{http://dx.doi.org/10.1109/TIE.2014.2316234}{DOI
  10.1109/TIE.2014.2316234}, no.~2, pp. 1270--1278, 2015.

\bibitem{DBLP:journals/iandc/MeseguerR13}
J.~Meseguer and G.~Rosu, ``The rewriting logic semantics project: A progress
  report,'' \emph{Inf. Comput.}, vol. 231,
  \href{http://dx.doi.org/10.1016/j.ic.2013.08.004}{DOI
  10.1016/j.ic.2013.08.004}, pp. 38--69, 2013.

\bibitem{DBLP:journals/cacm/Leroy09}
X.~Leroy, ``Formal verification of a realistic compiler,'' \emph{Commun.
  {ACM}}, vol.~52, \href{http://dx.doi.org/10.1145/1538788.1538814}{DOI
  10.1145/1538788.1538814}, no.~7, pp. 107--115, 2009.

\bibitem{DBLP:conf/sosp/KleinEHACDEEKNSTW09}
G.~Klein, K.~Elphinstone, G.~Heiser, J.~Andronick, D.~Cock, P.~Derrin,
  D.~Elkaduwe, K.~Engelhardt, R.~Kolanski, M.~Norrish, T.~Sewell, H.~Tuch, and
  S.~Winwood, ``sel4: formal verification of an {OS} kernel,'' in
  \emph{Proceedings of the 22nd {ACM} Symposium on Operating Systems Principles
  2009}, \href{http://dx.doi.org/10.1145/1629575.1629596}{DOI
  10.1145/1629575.1629596}, 2009, pp. 207--220.

\bibitem{TianD2011}
C.~Tian and Z.~Duan, ``Model checking rate monotonic scheduling algorithm based
  on propositional projection temporal logic,'' \emph{Journal of Software},
  vol.~22, no.~2, pp. 211--221, 2011, in {Chinese}.

\bibitem{DBLP:conf/iceccs/CuiDT14}
J.~Cui, Z.~Duan, and C.~Tian, ``Model checking rate-monotonic scheduler with
  {TMSVL},'' in \emph{19th International Conference on Engineering of Complex
  Computer Systems}, \href{http://dx.doi.org/10.1109/ICECCS.2014.37}{DOI
  10.1109/ICECCS.2014.37}, 2014, pp. 202--205.

\bibitem{DBLP:journals/lisp/OlveczkyM07}
P.~C. {\"{O}}lveczky and J.~Meseguer, ``Semantics and pragmatics of {Real-Time
  Maude},'' \emph{Higher-Order and Symbolic Computation}, vol.~20,
  \href{http://dx.doi.org/10.1007/s10990-007-9001-5}{DOI
  10.1007/s10990-007-9001-5}, no. 1-2, pp. 161--196, 2007.

\bibitem{DBLP:journals/tcs/ClavelDELMMQ02}
M.~Clavel, F.~Dur{\'{a}}n, S.~Eker, P.~Lincoln, N.~Mart{\'{\i}}{-}Oliet,
  J.~Meseguer, and J.~F. Quesada, ``Maude: specification and programming in
  rewriting logic,'' \emph{Theor. Comput. Sci.}, vol. 285,
  \href{http://dx.doi.org/10.1016/S0304-3975(01)00359-0}{DOI
  10.1016/S0304-3975(01)00359-0}, no.~2, pp. 187--243, 2002.

\bibitem{DBLP:journals/jlp/Meseguer12}
J.~Meseguer, ``Twenty years of rewriting logic,'' \emph{J. Log. Algebr.
  Program.}, vol.~81, \href{http://dx.doi.org/10.1016/j.jlap.2012.06.003}{DOI
  10.1016/j.jlap.2012.06.003}, no. 7-8, pp. 721--781, 2012.

\bibitem{DBLP:journals/entcs/OlveczkyM07a}
P.~C. {\"{O}}lveczky and J.~Meseguer, ``Abstraction and completeness for
  {Real-Time Maude},'' \emph{Electr. Notes Theor. Comput. Sci.}, vol. 176,
  \href{http://dx.doi.org/10.1016/j.entcs.2007.06.005}{DOI
  10.1016/j.entcs.2007.06.005}, no.~4, pp. 5--27, 2007.

\bibitem{DBLP:journals/tse/KatcherAS93}
D.~I. Katcher, H.~Arakawa, and J.~K. Strosnider, ``Engineering and analysis of
  fixed priority schedulers,'' \emph{{IEEE} Trans. Softw. Eng.}, vol.~19,
  \href{http://dx.doi.org/10.1109/32.241774}{DOI 10.1109/32.241774}, no.~9, pp.
  920--934, 1993.

\bibitem{DBLP:journals/tse/Holzmann97}
G.~J. Holzmann, ``The model checker {SPIN},'' \emph{{IEEE} Trans. Softw. Eng.},
  vol.~23, \href{http://dx.doi.org/10.1109/32.588521}{DOI 10.1109/32.588521},
  no.~5, pp. 279--295, 1997.

\bibitem{DBLP:conf/icfem/HanDW12}
M.~Han, Z.~Duan, and X.~Wang, ``Time constraints with temporal logic
  programming,'' in \emph{Formal Methods and Software Engineering - 14th
  International Conference on Formal Engineering Methods, {ICFEM} 2012.
  Proceedings}, ser. Lecture Notes in Computer Science, vol. 7635,
  \href{http://dx.doi.org/10.1007/978-3-642-34281-3\_20}{DOI
  10.1007/978-3-642-34281-3\_20}, 2012, pp. 266--282.

\end{thebibliography}\

%% Biographies

\begin{IEEEbiography}[{\includegraphics[width=1in,height=1.25in,clip,keepaspectratio]{AUTHOR_JIAXIANG_LIU.jpeg}}]%
{Jiaxiang Liu} received his B.S.\ degree in software engineering from
Tsinghua University, Beijing, China in 2010. Since 2010, he has been
working toward the Ph.D.\ degree in computer science at the Department
of Computer Science and Technology, Tsinghua University, Beijing,
China.

His research interests include formal methods, rewrite systems and
embedded systems.
\end{IEEEbiography}

\vspace{-5mm}
\begin{IEEEbiography}[{\includegraphics[width=1in,height=1.25in,clip,keepaspectratio]{AUTHOR_MIN_ZHOU.jpeg}}]%
{Min Zhou} received his B.S.\ degree in mathematics, and the
Ph.D.\ degree in computer science from Tsinghua University, Beijing,
China in 2007 and 2014, respectively.

He is currently working as a lecturer at the School of Software,
Tsinghua University, Beijing, China. His research interests include
model checking, program analysis and testing.
\end{IEEEbiography}

\vspace{-5mm}
\begin{IEEEbiography}[{\includegraphics[width=1in,height=1.25in,clip,keepaspectratio]{AUTHOR_XIAOYU_SONG.jpeg}}]%
{Xiaoyu Song} received the Ph.D.\ degree from the University of Pisa,
Italy, in 1991. From 1992 to 1998, he was on the faculty at the
University of Montreal, Canada. He joined the Department of Electrical
and Computer Engineering at Portland State University in 1998, where
he is now a professor. He was an editor of IEEE Transactions on VLSI
Systems and IEEE Transactions on Circuits and Systems. He was awarded
an Intel Faculty Fellowship from 2000 to 2005. His research interests
include formal methods, design automation, embedded systems and
emerging technologies.
\end{IEEEbiography}

\vspace{-5mm}
\begin{IEEEbiography}[{\includegraphics[width=1in,height=1.25in,clip,keepaspectratio]{AUTHOR_MING_GU.jpeg}}]%
{Ming Gu} received the B.S.\ degree in computer science from the
National University of Defense Technology, China in 1984, and the
M.S.\ degree in computer science from the Chinese Academy of Science,
China in 1986.

She is currently a professor at the School of Software, Tsinghua
University, China.  Her research interests include software formal
methods, software trustworthiness and middleware technology.
\end{IEEEbiography}

\vspace{-5mm}
\begin{IEEEbiography}[{\includegraphics[width=1in,height=1.25in,clip,keepaspectratio]{AUTHOR_JIAGUANG_SUN.jpeg}}]%
{Jiaguang Sun} received the B.S.\ degree in automation science from
Tsinghua University, China in 1970.

He is currently a professor and dean of the School of Information
Science and Technology in Tsinghua University, China. He is also a
member of the Chinese Academy of Engineering, and director of Tsinghua
National Laboratory for Information Science and Technology, China.
His research interests include computer graphics, computer-aided
design, formal verification of software, software engineering and
system architecture.
\end{IEEEbiography}
\end{document}